---
title: "Trade With Price Improvement"
author: "Bagas Trilaksonoaji"
date: "`r Sys.Date()`"
format: 
  html:
    toc: true
    number-sections: true
execute: 
  echo: false
  message: false
  warning: false
cache: TRUE
editor_options: 
  chunk_output_type: inline
---

<!-- link git repo:  -->

# Abstract 150

# Introduction 

There were numerous equities exchange established in the long history of Australia. Almost every big cities in Australia once had their own stock exchange, namely Sydney, Brisbane, Melbourne, Adelaide, Perth, even Hobart. In 1987, all six stock exchanges were merged under legislation of Australian Parliament and became Australian Stock Exchange [ASX-Story](https://www.asx.com.au/about/asx-story). Forward to two decades, this equities exchange merged with Sydney Futures Exchange to become Australian Securities Exchange. Now, its official name is ASX Limited, but it is easier to just cite its three leters: ASX. In this report, however, we will use the term ASX to refer the company or the equities market.

As a market operator, ASX main tasks include reviewing listing applications, supervising listed entities as well as listed brokers, operate and maintain electronic trading platform for buying and selling stocks, and clearing and settlement services. [ASX-About-Us](https://www.asx.com.au/about#:~:text=As%20an%20integrated%20exchange%2C%20ASX's,equities%20and%20fixed%20income%20markets.).

ASX operates under the regulatory framework of two government agencies, the Australia Securities and Investments Commission (ASIC) and the Reserve Bank of Australia (RBA) [ASX-Regulatory-Framework](https://www.asx.com.au/about/regulation). ASIC role is to supervise the stock market trading and also enforce its law related to financial market conduct, whereas RBA is responsible for supervising clearing and settlement related services.

## Background and Motivation 

Anyone who wants to buy or sell stocks can place an order on the ASX trading platform. Where the price of the buy and sell side is match, the trade can be executed. Any outstanding orders will still on the platform and are visible to anyone who have access to the market. This mechanism promote transparency and market equilibrium of supply and demand, hence ensure fair stock price formation [ASIC, 2022](https://asic.gov.au/about-asic/corporate-publications/newsletters/market-integrity-update/miu-issue-137-may-2022/). That is why we often call this trading venue provided by the ASX trading platform as `lit market` or `lit pool`. 

On the other hand, there are other venues that do not disclose the orders placed, at least until the transaction is executed. We will use the term `dark pool` to refer this kind of trading venue. There are several `dark pool` available for stock trading in Australia. However, in this project we will focus on the `dark pool` executed on the broker's internal crossing system. 

Crossing system is an electronic platform provided by the brokers to match orders placed by their clients with their other clients or with their own account (ASIC, 2014). On other words, orders can be matched on the broker's internal system before they are passed through to the ASX trading platform. 

What is the benefit for brokers and traders if they trades on the crossing system instead of ASX `lit pool`? One of the main reason is because the trading fees are much cheaper. Trades on the ASX platform are charged about 0.15 basis points per trade, whereas internal crossing charged as much as 0.04 basis points per trade, capped at $1,000 per month ([ASX Schedule Fees](https://asxonline.com/content/dam/asxonline/public/documents/schedule-of-fees/asx-trade-markets-participant-and-trading-schedule-of-fees.pdf)). The difference is quite huge given how much a broker can trade in daily basis.  

One important thing to note is that there was a difference in the internal crossing rule before and after 26 May 2013:

1. ***Before***
        
    Based on the Rule 4.2.3(1)(b) of the ASIC Market Integrity Rules (Competition in Exchange Markets) 2011, the trades done in the broker's internal crossing system are referred as `Trades At or Within The Spread`. As the name suggests, the executed trade price must be at the same as the best available bid or best available offer of the `lit market` (at-the-spread), or better (within-the-spread). Meaning that the price executed on this `dark pool` cannot be less than the best bid price or more than the best offer price in the `lit market`. 
2. ***After***
    
    The rule had changed, based on the Rule 4.2.3 of the ASIC Market Integrity Rules (Competition in Exchange Markets) Amendment 2012 (No. 1), the trades on the crossing system must provide meaningful price improvement (ASIC, 2014). Trades at the best bid or best offer price of the `lit market` is not in the option anymore. This rule was effective on 26 May 2013. ASIC and ASX refer this internal crossing trade as `Trades With Price Improvement`, or next we can refer this as TWPI. This report will focus on this kind of dark trades.

The amendment of the rule in May 2013 was to address the negative impact of the dark trades on the price formation, specifically from the `at-the-spread` trades (ASIC, 2014), due to traders may prefer to do the trades on the `dark pool` as it has less fees compared to `lit pool`. The adverse effect on the price discovery was one of the main reason why the regulator wants more trading to occur on the `lit pool`.

On the other side of point of view, there was also economic incentive of this to the ASX. On  14 October 1998, ASX become a public-owned companies after demutualising itself. The main analysis behind this was that "***ASX needed to become more flexible, responsive and commercially focused, capable of quickly taking up emerging commercial opportunities***" (ASIC, 2004). Therefore, as an all for-profit enterprise, ASX would take any business opportunity to raise their source of income to maximize their returns to shareholders. ASIC amended the rules in order to provide meaningful price improvement would move a significant portion of the dark trades to the ASX platform, which in turn rise its profit from trading fees.

**Key Motivation**

On May 2022, ASIC release a Market Integrity Update (Issue 137) article reviewing TWPI. In the article, ASIC mentioned that several participants were reporting a significant number of their TWPIs without price improvement. ASIC emphasized that such trades must be done on `lit market` because it will promote liquidity and price formation. TWPI is an exception because it offers better prices for the investor, the price that the `lit market` cannot provide. 

Following up this Market Integrity Update, ASIC issued an infringement notice to Wilson Advisory and Stockbroking Pty Ltd to pay to ASIC, behalf of the Commonwealth, in the sum of $548,328. The notice given on 16 December 2022, while published on ASIC media release on 3 February 2023. 

The penalty shows that ASIC has started to taken into account the breach of TWPI rule more seriously. However, since the rule was long introduced (in 2011 and amended in 2013), but only occurred to be ASIC concern until recently, I am intrigued to understand what happened to the TWPI just after the rule amendment in 26 May 2013. Were there no violation on the TWPI rule in its early establishment? Or there were other reasons that regulator had not put serious concern towards this issue? 

## Purpose and Key Questions 500

- understand how the rule change in 2013 impact at-spread crossing (percentage of at spread crossing to all twpi before and after rule change)
- how the rule change in 2013 moves the at-spread crossings to lit market (percentage of total twpi to total on market trade before and after rule change, also percentage of within-spread, at-spread, outside-spread, to total market trade)
- understand the violations after the rule change, how many (in dollar value), who did it, and the pattern (develop a model for finding the pattern)

Key Questions:
- any significant decrease in the percentage of at-spread crossings after the rule change?
- significant decrease in the percentage of TWPI after the rule change? meaning the twpi move to lit market
- how many violations after the rule change (in number and dollar value)
- who did the most violations? what kind of brokers?
- what conditions contributed to the occurrence of the violations?


- Brief information about method and data

## Significance and Contributions 500

- economic significance: how twpi violations contribute xxx dollar transaction. how many fees should be paid to market operator (ASX) 
- policy/regulatory significance: do asx need to address the policy of twpi reporting? now, reported twpi cannot be rejected. give them the pattern or model, to predict violation
- theoretical significance: give contribution to the academic understanding on what variables contributed to the occurrence of twpi violations

Novel features:
- new study related to twpi violation. before, many paper talks about special order and hidden orders
- the study includes detection model to help determine the violation pattern. never been used for twpi violation prediction

# Data and Methodology 1000

# Result and Discussion 2000

```{r}
library(tidyverse)
library(hms)
library(knitr)
library(kableExtra)
options(knitr.table.html.attr = "quarto-disable-processing=true")
library(tidymodels)
library(discrim)
library(rpart)
library(rpart.plot)
library(randomForest)
library(themis)

select(-Count, -(6:11), -(14:16), -(19:21))
```

```{r, eval = FALSE}
# list all csv files
list_2012 <- list.files(path = "data/2012/", pattern = "*.csv", full.names = TRUE)
list_2013 <- list.files(path = "data/2013/", pattern = "*.csv", full.names = TRUE)
list_2014 <- list.files(path = "data/2014/", pattern = "*.csv", full.names = TRUE)
list_2015 <- list.files(path = "data/2015/", pattern = "*.csv", full.names = TRUE)

# read all CSV files into a list of data frames
df_list_2012 <- lapply(list_2012, function(file) {
  read.csv(file, colClasses = c(Mykey = "character", BidID = "character", AskID = "character"))
  })
df_list_2013 <- lapply(list_2013, function(file) {
  read.csv(file, colClasses = c(Mykey = "character", BidID = "character", AskID = "character"))
  })
df_list_2014 <- lapply(list_2014, function(file) {
  read.csv(file, colClasses = c(Mykey = "character", BidID = "character", AskID = "character"))
  })
df_list_2015 <- lapply(list_2015, function(file) {
  read.csv(file, colClasses = c(Mykey = "character", BidID = "character", AskID = "character"))
  })

df_list <- c(df_list_2012, df_list_2013, df_list_2014, df_list_2015)

# combine the data frames into a single data frame
twpi_df <- bind_rows(df_list)

# save the df
save(twpi_df, file = "data/twpi_2012-2015")

```

```{r}
load("data/twpi_2012-2015")
twpi4 <- twpi_df |>
  select(-Count, -(6:11), -(14:16), -(19:21)) |>
  mutate(spread = abs(Ask_before - Bid_before),
         RecordDate = lubridate::ymd(RecordDate),
         Hour = substr(HourMinuteSecond, 1, 2),
         Year = lubridate::year(RecordDate),
         Month = lubridate::month(RecordDate),
         Date = lubridate::day(RecordDate),
         Dayofweek = lubridate::wday(RecordDate)) |>
  select(1:3, 5:11, 13:18, Marks) |>
  mutate(BuyerBrokerID = as.factor(BuyerBrokerID),
         across(12:14, as.factor),
         across(16:17, as.factor)) |>
  group_by(Instrument) |>
  mutate_if(is.numeric, function(x) (x-mean(x, na.rm = TRUE))/sd(x, na.rm = TRUE)) |>
  ungroup()

twpi4 <- twpi4 |> # remove duplicates
  group_by(Mykey) |>
  filter(n() == 1) |>
  ungroup()
```

```{r}
twpi_bf <- twpi4 |>
  filter(RecordDate < lubridate::ymd("20130526"))

twpi_af <- twpi4 |>
  filter(RecordDate >= lubridate::ymd("20130526"))
```

```{r}
# load broker classification data

broker <- read.csv("data/BrokerClassifications.csv")
class <- read.csv("data/Classifications.csv")
broker2 <- broker |> left_join(class, by = "Classification") |>
  mutate(No = as.factor(No))
```

## Analysis

```{r}
# sort top 10 broker 

broker_10 <- twpi_af |> 
  count(BuyerBrokerID) |>
  arrange(desc(n)) |>
  head(10) |>
  pull(BuyerBrokerID)

twpi_af |> 
  filter(BuyerBrokerID %in% broker_10) |>
  group_by(Marks) |>
  count(BuyerBrokerID) |>
  mutate(percentage = round(n / sum(n) * 100), 2) |>
  ungroup() |>
  filter(Marks != 1) |>
  mutate(Marks = ifelse(Marks == 2, "At Spread", "Outside")) |>
  ggplot(aes(x = reorder(BuyerBrokerID, -n), y = percentage, fill = Marks)) +
  geom_col(position = "dodge") +
  labs(title = "Violation in Percentage to TWPI",
       x = "Broker ID",
       y = "Percentage to TWPI") +
  theme_bw()
```

```{r}
# prepare data to model, only top 2 brokers

data <- twpi_af |> 
  filter(BuyerBrokerID %in% c(210,361),
         Hour != 16) |>
  # left_join(broker2[, c("No", "Description")], by = c("BuyerBrokerID" = "No")) |>
  mutate(Violation = as.factor(ifelse(Marks == 1, 0, 1)),
         Violation_1 = as.factor(ifelse(Marks == 2, 1, 0)),
         Violation_2 = as.factor(ifelse(Marks == 3, 1, 0))) |>
  select(2, 6, 8:16, 18:20) |>
  na.omit()

viol_at <- data |> select(-12, -14) 
viol_out <- data |> select(-12, -13) 

set.seed(1010)
va_split <- initial_split(viol_at, 2/3, strata = Violation_1)
va_train <- training(va_split)
va_test <- testing(va_split)

vo_split <- initial_split(viol_out, 2/3, strata = Violation_2)
vo_train <- training(vo_split)
vo_test <- testing(vo_split)

```

### Logistic Regression

```{r, eval = FALSE}
log_mod <- logistic_reg() |>
  set_engine("glm") |>
  set_mode("classification") |>
  translate()
log_fit1 <- log_mod |>
  fit(Violation_1 ~ .,
      data = va_train[, -c(1, 8, 9)])
log_fit2 <- log_mod |>
  fit(Violation_2 ~ .,
      data = vo_train[, -c(1, 8, 9)])

tidy(log_fit1) 
glance(log_fit1)

tidy(log_fit2) 
glance(log_fit2)

saveRDS(log_fit1, file = "data/report/log_fit1.rds")
saveRDS(log_fit2, file = "data/report/log_fit2.rds")
```

Balance accuracy of at-the-spread violation on the test set:

```{r}
# measure accuracy logistic regression on at-spread violation 
log_fit1 <- readRDS("data/report/log_fit1.rds")

treshold <- 0.01625

va_tr_pred <- log_fit1 |> 
  augment(new_data = va_train) |>
  mutate(pred_viol = as.factor(ifelse(.pred_1 > treshold, 1, 0)))
va_ts_pred <- log_fit1 |> 
  augment(new_data = va_test) |>
  mutate(pred_viol = as.factor(ifelse(.pred_1 > treshold, 1, 0)))

matrix_va_tr <- va_tr_pred |>
  count(Violation_1, pred_viol) |>
  group_by(Violation_1) |>
  mutate(cl_acc = n[pred_viol == Violation_1]/sum(n)) |>
  pivot_wider(names_from = pred_viol, 
              values_from = n, values_fill=0)

matrix_va_ts <- va_ts_pred |>
  count(Violation_1, pred_viol) |>
  group_by(Violation_1) |>
  mutate(cl_acc = n[pred_viol == Violation_1]/sum(n)) |>
  pivot_wider(names_from = pred_viol, 
              values_from = n, values_fill=0)

bal_accuracy_logistic_va <- bal_accuracy(va_ts_pred, Violation_1, pred_viol)$.estimate
bal_accuracy_logistic_va
```

Balance accuracy of outside-the-spread violation on the test set:

```{r}
# measure accuracy logistic regression on outside violation 
log_fit2 <- readRDS("data/report/log_fit2.rds")

treshold <- 0.045

vo_tr_pred <- log_fit2 |> 
  augment(new_data = vo_train) |>
  mutate(pred_viol = as.factor(ifelse(.pred_1 > treshold, 1, 0)))
vo_ts_pred <- log_fit2 |> 
  augment(new_data = vo_test) |>
  mutate(pred_viol = as.factor(ifelse(.pred_1 > treshold, 1, 0)))

matrix_vo_tr <- vo_tr_pred |>
  count(Violation_2, pred_viol) |>
  group_by(Violation_2) |>
  mutate(cl_acc = n[pred_viol == Violation_2]/sum(n)) |>
  pivot_wider(names_from = pred_viol, 
              values_from = n, values_fill=0)

matrix_vo_ts <- vo_ts_pred |>
  count(Violation_2, pred_viol) |>
  group_by(Violation_2) |>
  mutate(cl_acc = n[pred_viol == Violation_2]/sum(n)) |>
  pivot_wider(names_from = pred_viol, 
              values_from = n, values_fill=0)


bal_accuracy_logistic_vo <- bal_accuracy(vo_ts_pred, Violation_2, pred_viol)$.estimate
bal_accuracy_logistic_vo
```

#### Apply model to twpi data outside broker 210 and 361

```{r}
# filter data
data_2 <- twpi_af |> 
  filter(!(BuyerBrokerID %in% c(210,361)),
         Hour != 16) |>
  # left_join(broker2[, c("No", "Description")], by = c("BuyerBrokerID" = "No")) |>
  mutate(Violation = as.factor(ifelse(Marks == 1, 0, 1)),
         Violation_1 = as.factor(ifelse(Marks == 2, 1, 0)),
         Violation_2 = as.factor(ifelse(Marks == 3, 1, 0))) |>
  select(2, 6, 8:16, 18:20) |>
  na.omit()
```

Balance accuracy of at-the-spread violation on other brokers TWPI:

```{r}
# apply at-spread violation model and measure accuracy 
treshold <- 0.01625

data_2_va_pred <- log_fit1 |> 
  augment(new_data = data_2) |>
  mutate(pred_viol = as.factor(ifelse(.pred_1 > treshold, 1, 0)))

matrix_data_2_va <- data_2_va_pred |>
  count(Violation_1, pred_viol) |>
  group_by(Violation_1) |>
  mutate(cl_acc = n[pred_viol == Violation_1]/sum(n)) |>
  pivot_wider(names_from = pred_viol, 
              values_from = n, values_fill=0)


bal_accuracy_logistic_data2_1 <- bal_accuracy(data_2_va_pred, Violation_1, pred_viol)$.estimate
bal_accuracy_logistic_data2_1
```

Balance accuracy of outside-the-spread violation on other brokers TWPI:

```{r}
# apply outside violation model and measure accuracy 
treshold <- 0.045

data_2_vo_pred <- log_fit2 |> 
  augment(new_data = data_2) |>
  mutate(pred_viol = as.factor(ifelse(.pred_1 > treshold, 1, 0)))

matrix_data_2_vo <- data_2_vo_pred |>
  count(Violation_2, pred_viol) |>
  group_by(Violation_2) |>
  mutate(cl_acc = n[pred_viol == Violation_2]/sum(n)) |>
  pivot_wider(names_from = pred_viol, 
              values_from = n, values_fill=0)


bal_accuracy_logistic_data2_2 <- bal_accuracy(data_2_vo_pred, Violation_2, pred_viol)$.estimate
bal_accuracy_logistic_data2_2
```

### Random Forest

#### At-The-Spread

```{r}
data_rec1 <- recipe(Violation_1 ~ ., data = va_train) |>
  step_downsample(Violation_1) |>
  step_rm(Instrument, Year, Month)
```

```{r}
# data_prep1 <- prep(data_rec1)
# saveRDS(data_prep1, file = "data/report/data_prep1.rds")

data_prep1 <- readRDS("data/report/data_prep1.rds")
data_juice1 <- juice(data_prep1)
```

```{r, eval = FALSE}
# tuning hyperparameters
rf_tune1 <- rand_forest(
  mtry = tune(),
  trees = 1000,
  min_n = tune()) |>
  set_mode("classification") |>
  set_engine("randomForest")

tune_wf1 <- workflow() |>
  add_recipe(data_rec1) |>
  add_model(rf_tune1)
  
doParallel::registerDoParallel() # for parallel processing

set.seed(3344)
w_folds1 <- vfold_cv(va_train)

rf_grid1 <- grid_regular(mtry(range = c(2, 8)),
                        min_n(range = c(1, 4)),
                        levels = 3)

tune_res1 <- 
  tune_grid(tune_wf1,
            resamples = w_folds1,
            grid = rf_grid1
            )

# select best combination based on accuracy
best_hp1 <- tune_res1 |> select_best(metric = "accuracy")
```

```{r, eval = FALSE}
# fit the model
set.seed(3344)
rf1 <- rand_forest(mtry = 5, 
                     min_n = 4) |>
  set_mode("classification") |>
  set_engine("randomForest", importance = TRUE)

rf_workflow1 <- workflow() |>
  add_model(rf1) |>
  add_formula(Violation_1 ~ .)

fit_rf1 <- rf_workflow1 |> 
  fit(data = data_juice1)

saveRDS(fit_rf1, file = "data/report/fit_rf1.rds")
```

```{r}
# apply the model to test set
fit_rf1 <- readRDS("data/pattern/fit_rf1.rds")

va_ts_pred_rf <- bake(data_prep1, va_test) |>
  mutate(pviol = predict(fit_rf1,
                         bake(data_prep1, va_test))$.pred_class)
```

Balance accuracy of at-the-spread violation on the test set, and the variable importance:

```{r}
# measure accuracy

cf_rf_ts1 <- va_ts_pred_rf |>
  count(Violation_1, pviol) |>
  group_by(Violation_1) |>
  mutate(Accuracy = n[Violation_1==pviol]/sum(n)) |>
  pivot_wider(names_from = "pviol", 
              values_from = n, values_fill = 0)


bal_accuracy_rf1 <- bal_accuracy(va_ts_pred_rf, Violation_1, pviol)$.estimate

bal_accuracy_rf1

rf_model1 <- pull_workflow_fit(fit_rf1)$fit
importance(rf_model1)
```

#### Outside the Spread

```{r}
data_rec2 <- recipe(Violation_2 ~ ., data = vo_train) |>
  step_downsample(Violation_2) |>
  step_rm(Instrument, Year, Month)
```

```{r}
# data_prep2 <- prep(data_rec2)
# saveRDS(data_prep2, file = "data/report/data_prep2.rds")

data_prep2 <- readRDS("data/report/data_prep2.rds")
data_juice2 <- juice(data_prep2)
```

```{r, eval = FALSE}
# tuning hyperparameters
rf_tune2 <- rand_forest(
  mtry = tune(),
  trees = 1000,
  min_n = tune()) |>
  set_mode("classification") |>
  set_engine("randomForest")

tune_wf2 <- workflow() |>
  add_recipe(data_rec2) |>
  add_model(rf_tune2)
  
doParallel::registerDoParallel() # for parallel processing

set.seed(3344)
w_folds2 <- vfold_cv(vo_train)

rf_grid2 <- grid_regular(mtry(range = c(2, 8)),
                        min_n(range = c(1, 4)),
                        levels = 3)

tune_res2 <- 
  tune_grid(tune_wf2,
            resamples = w_folds2,
            grid = rf_grid2
            )

# select best combination based on accuracy
best_hp2 <- tune_res2 |> select_best(metric = "accuracy")
```

```{r, eval = FALSE}
# fit the model
set.seed(3344)
rf2 <- rand_forest(mtry = 5, 
                     min_n = 4) |>
  set_mode("classification") |>
  set_engine("randomForest", importance = TRUE)

rf_workflow2 <- workflow() |>
  add_model(rf2) |>
  add_formula(Violation_2 ~ .)

fit_rf2 <- rf_workflow2 |> 
  fit(data = data_juice2)

saveRDS(fit_rf2, file = "data/report/fit_rf2.rds")
```

```{r}
# apply the model to test set
fit_rf2 <- readRDS("data/report/fit_rf2.rds")

vo_ts_pred_rf <- bake(data_prep2, vo_test) |>
  mutate(pviol = predict(fit_rf2,
                         bake(data_prep2, vo_test))$.pred_class)
```

Balance accuracy of outside-the-spread violation on the test set, and the variable importance:

```{r}
# measure accuracy
cf_rf_ts2 <- vo_ts_pred_rf |>
  count(Violation_2, pviol) |>
  group_by(Violation_2) |>
  mutate(Accuracy = n[Violation_2==pviol]/sum(n)) |>
  pivot_wider(names_from = "pviol", 
              values_from = n, values_fill = 0)

bal_accuracy_rf2 <- bal_accuracy(vo_ts_pred_rf, Violation_2, pviol)$.estimate

bal_accuracy_rf2

rf_model2 <- pull_workflow_fit(fit_rf2)$fit
importance(rf_model2)
```

#### Apply model to twpi data outside broker 210 and 361

Balance accuracy of at-the-spread violation on the other brokers TWPI:

```{r}
# apply at-spread violation model and measure accuracy 
data_2_va_pred_rf <- bake(data_prep1, data_2) |>
  mutate(pviol = predict(fit_rf1,
                         bake(data_prep1, data_2))$.pred_class)

cf_rf_data_2_1 <- data_2_va_pred_rf |>
  count(Violation_1, pviol) |>
  group_by(Violation_1) |>
  mutate(Accuracy = n[Violation_1==pviol]/sum(n)) |>
  pivot_wider(names_from = "pviol", 
              values_from = n, values_fill = 0)

bal_accuracy_rf_data2_1 <- bal_accuracy(data_2_va_pred_rf, Violation_1, pviol)$.estimate
bal_accuracy_rf_data2_1
```

Balance accuracy of outside-the-spread violation on the other brokers TWPI:

```{r}
# apply outside violation model and measure accuracy 
data_2_vo_pred_rf <- bake(data_prep2, data_2) |>
  mutate(pviol = predict(fit_rf2,
                         bake(data_prep2, data_2))$.pred_class)

cf_rf_data_2_2 <- data_2_vo_pred_rf |>
  count(Violation_2, pviol) |>
  group_by(Violation_2) |>
  mutate(Accuracy = n[Violation_2==pviol]/sum(n)) |>
  pivot_wider(names_from = "pviol", 
              values_from = n, values_fill = 0)

bal_accuracy_rf_data2_2 <- bal_accuracy(data_2_vo_pred_rf, Violation_2, pviol)$.estimate
bal_accuracy_rf_data2_2
```

### Summary

```{r}
#| tbl-colwidths: 3

summary <- data.frame(
  Violation = c("at-spread", "outside-spread"),
  GLM_top2 = c(bal_accuracy_logistic_va, bal_accuracy_logistic_vo),
  GLM_others = c(bal_accuracy_logistic_data2_1, bal_accuracy_logistic_data2_2),
  RF_top2 = c(bal_accuracy_rf1, bal_accuracy_rf2),
  RF_others = c(bal_accuracy_rf_data2_1, bal_accuracy_rf_data2_2)
)


summary |>
  kbl() |>
  kable_styling(full_width = FALSE, position = "left")
```


### Limitation

# Conclusion 500

## Potential Future Work

# References

ASX-Story https://www.asx.com.au/about/asx-story

ASX-About Us https://www.asx.com.au/about#:~:text=As%20an%20integrated%20exchange%2C%20ASX's,equities%20and%20fixed%20income%20markets.

ASX-About-Regulation https://www.asx.com.au/about/regulation

[ASIC, 2022](https://asic.gov.au/about-asic/corporate-publications/newsletters/market-integrity-update/miu-issue-137-may-2022/)

ASIC Review of recent rule changes affecting dark liquidity, report 394, may 2014

ASIC Market demutualisation and privatisation: The Australian experience, 20 may 2004

ASX Trade: Markets Participant and Trading Schedule of Fees https://asxonline.com/content/dam/asxonline/public/documents/schedule-of-fees/asx-trade-markets-participant-and-trading-schedule-of-fees.pdf

Market Integrity Update (Issue 137) https://asic.gov.au/about-asic/corporate-publications/newsletters/market-integrity-update/miu-issue-137-may-2022/#review-of-trades-with-price-improvement

wilson advisory infringement notice https://asic.gov.au/about-asic/news-centre/find-a-media-release/2023-releases/23-016mr-wilsons-advisory-and-stockbroking-ltd-acn-010-529-665-pays-548-328-infringement-notice/

