---
title: "Trade With Price Improvement"
author: "Bagas Trilaksonoaji"
date: "`r Sys.Date()`"
format: 
  pdf:
    toc: true
    number-sections: true
execute: 
  eval: TRUE
  echo: false
  message: false
  warning: false
cache: TRUE
editor_options: 
  chunk_output_type: inline
---

\pagebreak

# Abstract

# Introduction 

There were numerous equities exchanges established in the long history of Australia. Almost every big city in Australia once had their own stock exchange, namely Sydney, Brisbane, Melbourne, Adelaide, Perth, and even Hobart. The Australian Parliament merged all six stock exchanges in 1987, creating the Australian Stock Exchange [(Australian Stock Exchange, n.d.)](https://www.asx.com.au/about/asx-story). Two decades later, this equities exchange merged with Sydney Futures Exchange to become the Australian Securities Exchange. Although its official name is now ASX Limited, it is more convenient to simply refer to it as ASX. In this report, however, we will use the term ASX to refer to the company or the equities market.

As a market operator, ASX's main tasks include reviewing listing applications, supervising listed entities as well as listed brokers, operating and maintaining an electronic trading platform for buying and selling stocks, and providing clearing and settlement services. [(Australian Stock Exchange, n.d.)](https://www.asx.com.au/about#:~:text=As%20an%20integrated%20exchange%2C%20ASX's,equities%20and%20fixed%20income%20markets.).

ASX operates under the regulatory framework of two government agencies, the Australia Securities and Investments Commission (ASIC) and the Reserve Bank of Australia (RBA) [(Australian Stock Exchange, n.d.)](https://www.asx.com.au/about/regulation). ASIC oversees stock market trading and enforces financial market conduct laws, while the Reserve Bank of Australia (RBA) oversees clearing and settlement-related services.

## Background and Motivation 

Anyone who wants to buy or sell stocks can place an order on the ASX trading platform. When the price of the buy and sell side matches, the trade can be executed. Outstanding orders will remain on the platform and visible to market participants. This mechanism promotes transparency and market equilibrium in supply and demand, thereby ensuring fair stock price formation [(Australian Securities and Investments Commission, 2022)](https://asic.gov.au/about-asic/corporate-publications/newsletters/market-integrity-update/miu-issue-137-may-2022/). That is why we often call this trading venue provided by the ASX trading platform `lit market` or `lit pool`.

On the other hand, there are other venues that do not disclose the orders placed, at least until the transaction is executed. We will use the term `dark pool` to refer to this kind of trading venue. There are several `dark pools` available for stock trading in Australia. However, in this project, we will focus on the `dark pool` executed on the broker's internal crossing system. This kind of `dark pool` is often called NBBO trade, which stands for National Best Bid and Offer Prices ([Charles Lane Advisory Pty Ltd, 2014](https://download.asic.gov.au/media/1344590/rep394-attachment-published-19-May-2014.pdf)). 

Crossing system is an electronic platform provided by the brokers to match orders placed by their clients with their other clients or with their own account [(Australian Securities and Investments Commission, 2014)](https://download.asic.gov.au/media/1344596/rep394-published-19-May-2014.pdf). In other words, orders can be matched on the broker's internal system before they are passed through to the ASX trading platform. For the next section, we will use the term `crossing` to refer to the trades in the broker's internal crossing system.

What is the benefit for brokers and traders if they trade on the crossing system instead of the ASX `lit pool`? One of the primary benefits is the significantly lower trading fees. Trades on the ASX platform are charged about 0.15 basis points per trade, whereas internal crossing charges as much as 0.04 basis points per trade, capped at $1,000 per month [(Australian Securities Exchange, 2024)](https://asxonline.com/content/dam/asxonline/public/documents/schedule-of-fees/asx-trade-markets-participant-and-trading-schedule-of-fees.pdf)). The difference is quite huge given how much a broker can trade on a daily basis.

It is important to note that the rule in which internal crossings referred to was differed before and after 26 May 2013:

1. **Before**

    Rule 4.2.3(1)(b) of the ASIC Market Integrity Rules (Competition in Exchange Markets) 2011, regulates the `At or Within the Spread` trades. As the name suggests, the executed trade price must be at the same as the best available bid or best available offer of the `lit market` (at-the-spread), or better (within-the-spread). This implies that the price executed in this "dark pool" cannot be lower than the best bid price or higher than the best offer price in the "lit market".
2. **After**

    The rule had changed, based on Rule 4.2.3 of the ASIC Market Integrity Rules (Competition in Exchange Markets) Amendment 2012 (No. 1), the trades on the crossing system must provide meaningful price improvement [(Australian Securities and Investments Commission, 2014)](https://download.asic.gov.au/media/1344596/rep394-published-19-May-2014.pdf). Trades at the best bid or best offer price of the `lit market` are not in the option anymore. This rule was effective on 26 May 2013. ASIC and ASX refer to this rule as `Trades With Price Improvement`, or next we can refer to this as TWPI. This study will specifically focus on the internal crossings, which follow the TWPI rule.

The amendment of the rule in May 2013 was to address the negative impact of the dark trades on the price formation, specifically from the `at-the-spread` trades [(Australian Securities and Investments Commission, 2014)](https://download.asic.gov.au/media/1344596/rep394-published-19-May-2014.pdf), due to traders may prefer to do the trades on the `dark pool` as it costs less fees compared to the `lit pool`. The adverse effect on price discovery was one of the main reasons why the regulator wanted more trading to occur in the `lit pool`.

On the other hand, the ASX was also motivated by an economic incentive. On October 14, 1998, ASX became a publicly owned company after demutualising itself. The main analysis behind this was that "*ASX needed to become more flexible, responsive, and commercially focused, capable of quickly taking up emerging commercial opportunities*" [(Australian Securities and Investments Commission, 2004)](https://asic.gov.au/about-asic/news-centre/speeches/market-demutualisation-and-privatisation-the-australian-experience/). Therefore, ASX, being an all-for-profit enterprise, would seize any business opportunity to increase its revenue and maximize its returns to shareholders. ASIC amended the rules to provide meaningful price improvements, which would shift a significant portion of dark trades to the ASX platform, thereby increasing its profit from trading fees.

**Key Motivation**

In May 2022, ASIC published a Market Integrity Update (Issue 137), which included a review of TWPI. In the article, ASIC mentioned that several participants were reporting a significant number of their TWPIs without any price improvement. ASIC emphasized that such trades must be done in the `lit market` because they will promote liquidity and price formation. TWPI is an exception because it offers better prices for the investor—a price that the `lit market` cannot provide.

Following up on the Market Integrity Update, ASIC issued an infringement notice to Wilson Advisory and Stockbroking Pty Ltd, requiring them to pay $548,328 to ASIC on behalf of the Commonwealth. The notice was given on 16 December 2022, while published on an ASIC media release on 3 February 2023.

The penalty shows that ASIC has started to take the breach of the TWPI rule more seriously. Probably due to the increasing number of dark trades that should be done on the `lit market`, as ASIC intention of rule change in 2013 was to increasing the proportion of trading on the `lit market`, hence, enhancing the fairness on the market price formation [Australian Securities and Investments Commission, 2014](https://download.asic.gov.au/media/1344596/rep394-published-19-May-2014.pdf)

Based on these insights, I'm intrigued to analyse the `at-the-spread` dark trades before and after the rule amendment in 26 May 2013. Different from the ASIC report ([Report 394](https://download.asic.gov.au/media/1344596/rep394-published-19-May-2014.pdf)) on May 2014, this study will focus on the internal crossings mechanism, among all other dark pool mechanisms. Furthermore, the internal crossings violation on the TWPI rule will be analysed, along with the model that explains the pattern of the violation.

## Purpose and Key Questions

The main objective of this study is to look into how the rule change in 2013 affected `at-the-spread` crossings reported as TWPI. The way brokers act and trade on the `dark pool` should be greatly affected by this change, as they must move the `at-the-spread` crossings to the `lit pool`. Moreover, this study will also examine the violation of the TWPI after the rule change.

The study will compare transaction data before and after the 2013 regulation change, with a focus on `at-the-spread` crossings. Trade volumes, prices, time stamps, broker IDs, and several other variables will be included in the analysis. There will be simple statistical methods used to compare trade distributions, and machine learning methods may be used to find trends of the violations in the TWPI rule.

### Understand the Impact on At-the-Spread Crossings

`At-the-spread` crossings were possible before the rule change, which meant that trades could be made at the best bid or best offer price available on the `lit market`. However, the 2013 amendment says that trades must lead to price improvements, in other words, `at-the-spread` crossings are no longer allowed. The main question to ask is the proportion of `at-the-spread` crossings among all TWPI-reported trades before and after 26 May 2013. This comparison will help to figure out whether the rule change stopped people from doing `at the spread` crossings at the `dark pool`. Furthermore, the dollar value that possibly moved to the `lit market` from the `dark pool` will also be analysed.

**Key Question**:

1. Did the number of `at-the-spread` crossings drop significantly after 26 May 2013?
2. How much was approximately the value of `at-the-spread` internal crossings that moved to the lit market after the rule change?

### Identify the Violation Pattern

ASIC's media release and publication in the last couple of years suggested that there have been TWPI rule violations in recent years. This study will examine the monetary worth of the violations, identify which brokers do the most violations, and find any pattern by creating a model to help spot non-compliance in the future.

**Key Question**:

1. How many violations, both in terms of the quantity of trades and the dollar value, occurred following the rule change?
2. What kind of brokers violate the rules the most? Are some brokers more likely to break TWPI rules than others?
3. What conditions contributed to the occurrence of the violations?

## Significance and Contributions 500

There are three main areas in which this study may contribute: academic, regulatory, and economic.

### Economic Significance

The purpose of the rule amendment in 2013 was to encourage meaningful price improvement on the `dark pool` ([ASIC, 2014](https://download.asic.gov.au/media/1344596/rep394-published-19-May-2014.pdf)). Without price improvement, trading must be done on the ASX trading platform to promote liquidity on the market. Charles Lane Advisory Pty Ltd report ([Attachment to ASIC Report 394](https://download.asic.gov.au/media/1344590/rep394-attachment-published-19-May-2014.pdf)) in May 2014 has thoroughly analyse the impact of this rule change and stated that there was a decline in the dollar value from \$54.9 billion to \$21 billion, 100 days prior the rule change and after. Our study aims to validate this figure by focusing only on internal crossings and extending the analysis over a one-year period, with the goal of providing a more in-depth understanding of the rule's economic impacts.

From a financial point of view, it is also essential to comprehend TWPI violations since they have the potential to compromise market integrity in general [(Australian Securities and Investments Commission, 2014)](https://download.asic.gov.au/media/1344596/rep394-published-19-May-2014.pdf). By figuring out how much value in dollar TWPI violations occurred, this study can show how much value was traded not in a legal manner. If these trades had taken place on the `lit market`, then there would be trading fees that should be paid to the market operator (ASX). Internal crossing trades incur a fee of approximately 0.04 basis points per trade, whereas trades on the ASX platform cost 0.15 basis points per trade [(Australian Securities and Investments Commission, 2024)](https://asxonline.com/content/dam/asxonline/public/documents/schedule-of-fees/asx-trade-markets-participant-and-trading-schedule-of-fees.pdf). This significant difference in trading fees indicates additional economic implications.

### Regulatory Significance

Understanding the magnitude of the TWPI violation since its amendment in 2013 may provide regulators with new insights to address the TWPI reporting issues at the prevention level rather than through corrective action. Furthermore, if the detection model developed in this study proves effective, regulators might utilize it as part of an automated enforcement mechanism to ensure TWPI reporting follows the regulations.

### Academic Significance

There is a lot of academic literature on special orders and hidden orders, but this study adds a new dimension by focusing on TWPI trades. This study's novel contribution is the use of a prediction model to identify patterns and conditions that result in TWPI rule violations. The study might help academics learn more about how rules are enforced in financial markets and the situations where traders are most likely to take advantage of the system loopholes.

# Data and Methodology

This study use transaction and market data provided by The Securities Industry Research Center of Asia-Pacific (SIRCA). Founded in 1997, SIRCA is a non-profit organization mostly consisting of academics from Australian and New Zealand universities. It contains a variety of financial market data, including stock exchanges data, from Australia, New Zealand, and even the United States ([SIRCA, n.d.](https://www.sirca.org.au/about-sirca/)). SIRCA members are free to gather data for research purposes, provided that the university they are at have paid subscription fee to SIRCA.

The ASX transaction data provided by SIRCA consists of the stock code, date and time, price, volume, dollar value, broker ID, and other related variables. In addition, the ASX market data includes the best bid and best offer order prices, as well as the volume of those orders. These best bid and best offer prices are used to calculate the spread, price and volume, at each particular time.

We use datasets from stocks listed on the S&P/ASX200 index, as this index represents a group of stocks with the biggest market capitalization. This indicates that these stocks are the biggest and most valuable ones on the ASX. Market capitalization and liquidity often have a positive correlation. These traits from the S&P/ASX200 index ensure sufficient trading interest from major brokers (Duong et al., 2018), which is important for gauging the effect of the regulation change.

We will use data from the introduction of the `at or within the spread` rule in November 2011. However, since our investigation found that there were no internal crossings in 2011, we will use data from 2012 until the end of 2015.

The transaction dataset we obtained from SIRCA was filtered to include only internal crossing trades, marked as "OFFTR" in the `RecordType` variable and as "NX XT" in the `Qualifiers` variable. We combined this data with market data, allowing us to include every transactions with the best bid and offer price and volume of the market. Combining these many data requires enormous processing power, therefore, we use the help of Nectar Research Cloud, a cloud computing service from the Australia Research Data Commons.

**Methodology**

This study uses a methodical approach to look at how changes in regulations affect `at-the-spread` crossings in dark pools, particularly on the broker's internal crossing system. This study also looks for trends in TWPI rule violations. With the help of SIRCA's ASX transaction and market data, we hope to comprehend the regulatory and economic effects of the 2013 rule change.

*Rule Change Impact on At-the-Spread Crossings*

Analyzing the proportion before and after the rule adjustment on 26 May 2013 is the first step in determining how the rule change would affect `at-the-spread` transactions. The frequency and characteristics of `at-the-spread` crossings in the internal crossing system before and after the rule change will be compared. In order to guarantee sufficient liquidity and representation of the whole market, we concentrate our analysis on the S&P/ASX200 stocks. This allows us to determine if the regulatory adjustment was successful in limiting `at-the-spread` trades to close to none.

To assess the economic implications, we evaluate the aggregate dollar value of `at-the-spread` trades over two periods: the 100 trading days preceding and the 100 trading days succeeding the rule amendment on 26 May 2013. This controlled period reflects similar transaction trends or patterns. The value-based analysis will measure the economic consequences of trades that may have shifted to the 'lit market' following the amendment, thus assessing the possible enhancement in market transparency and fairness.

*Violation Pattern Analysis*

We utilize logistic regression and random forest models to discover and analyze violations of TWPI rules, as both are frequently employed for binary classification and pattern recognition. These models will assist in identifying circumstances in which TWPI violations are more likely. Every trade will be evaluated for compliance to the TWPI regulations, utilizing parameters such as trade volume, timing, and spread to discern the pattern. Logistic regression yields probabilistic estimates of violations, but random forest facilitates the examination of complex variable interactions, providing both interpretability and robustness.

*Statistical Measures and Tools*

We will evaluate model performance utilizing balanced accuracy, specificity, and sensitivity, especially considering the apparent imbalance between complying and non-compliant trades. `R` machine learning and statistical packages, such as `tidymodels` and `caret`, will be utilized for processing, modeling, and validating outcomes in implementation and classification analysis.

# Result and Discussion

```{r}
library(tidyverse)
library(hms)
library(knitr)
library(kableExtra)
options(knitr.table.html.attr = "quarto-disable-processing=true")
library(scales)
library(viridis)
library(tidymodels)
library(discrim)
library(rpart)
library(rpart.plot)
library(randomForest)
library(themis)
library(caret)
library(pROC)

```

```{r, eval = FALSE}
# list all csv files
list_2012 <- list.files(path = "data/2012/", pattern = "*.csv", full.names = TRUE)
list_2013 <- list.files(path = "data/2013/", pattern = "*.csv", full.names = TRUE)
list_2014 <- list.files(path = "data/2014/", pattern = "*.csv", full.names = TRUE)
list_2015 <- list.files(path = "data/2015/", pattern = "*.csv", full.names = TRUE)

# read all CSV files into a list of data frames
df_list_2012 <- lapply(list_2012, function(file) {
  read.csv(file, colClasses = c(Mykey = "character", BidID = "character", AskID = "character"))
  })
df_list_2013 <- lapply(list_2013, function(file) {
  read.csv(file, colClasses = c(Mykey = "character", BidID = "character", AskID = "character"))
  })
df_list_2014 <- lapply(list_2014, function(file) {
  read.csv(file, colClasses = c(Mykey = "character", BidID = "character", AskID = "character"))
  })
df_list_2015 <- lapply(list_2015, function(file) {
  read.csv(file, colClasses = c(Mykey = "character", BidID = "character", AskID = "character"))
  })

df_list <- c(df_list_2012, df_list_2013, df_list_2014, df_list_2015)

# combine the data frames into a single data frame
twpi_df <- bind_rows(df_list)

# save the df
save(twpi_df, file = "data/twpi_2012-2015")

```

```{r, eval = FALSE}
load("data/twpi_2012-2015")
twpi4 <- twpi_df |>
  select(-Count, -(6:11), -(14:16), -(19:21)) |>
  mutate(spread = abs(Ask_before - Bid_before),
         RecordDate = lubridate::ymd(RecordDate),
         Hour = substr(HourMinuteSecond, 1, 2),
         Year = lubridate::year(RecordDate),
         Month = lubridate::month(RecordDate),
         Date = lubridate::day(RecordDate),
         Dayofweek = lubridate::wday(RecordDate)) |>
  select(1:3, 5:11, 13:18, Marks) |>
  mutate(BuyerBrokerID = as.factor(BuyerBrokerID),
         across(12:14, as.factor),
         across(16:17, as.factor)) |>
  group_by(Instrument) |>
  mutate_if(is.numeric, function(x) (x-mean(x, na.rm = TRUE))/sd(x, na.rm = TRUE)) |>
  ungroup()

twpi4 <- twpi4 |> # remove duplicates
  group_by(Mykey) |>
  filter(n() == 1) |>
  ungroup()

save(twpi4, file = "data/report/twpi4")
```

```{r, eval = FALSE}
load("data/report/twpi4")

twpi_bf <- twpi4 |>
  filter(RecordDate < lubridate::ymd("20130526"))

twpi_af <- twpi4 |>
  filter(RecordDate >= lubridate::ymd("20130526"))
```

```{r, eval = FALSE}
# load broker classification data

broker <- read.csv("data/BrokerClassifications.csv")
class <- read.csv("data/Classifications.csv")
broker2 <- broker |> left_join(class, by = "Classification") |>
  mutate(No = as.factor(No))
```

## Rule Change Impact on At-the-Spread Crossings

We compare the number of internal crossings before and after the rule change in 26 May 2013. The `at-the-spread` internal crossings can be done before the meaningful price improvement rule introduction, hence, only `within-the-spread` crossings can be done after 26 May 2013. @tbl-bf and @tbl-af show monthly number and proportion of internal crossings before and after the rule change, respectively. In addition, the total proportion for each trade price group (within/at/outside-the-spread) can be seen from @tbl-bfaf.

The monthly proportion of `at-the-spread` internal crossings consistent around between 40% - 60% from the start of the rule until the amendment. The highest proportions were at October - December 2012, with 61% - 62% of the total internal crossings. The lowest proportion after taking out January 2012 (lowest number of crossings because of the transition) was at April 2013, with only 39% of the total crossings.

@tbl-bf shows that there was a number of violation in the early rule of internal crossings. However, the proportion were not significant, under 2% all the time except in September 2012. The total proportion of `at-the-spread` crossings from the introduction of the `at or within-the-spread` rule until the amendment in May 2013 was 52%.

These number and percentage shows that there were balanced proportion of `within-the-spread` and `at-the-spread` crossings, with only insignificant number of violation (`outside-the-spread`), from the introduction of the rule in November 2011 until the amendment in May 2013.

```{r}
#| label: tbl-bf
#| tbl-cap: "The Number and Proportion of Internal Crossings Before 26 May 2013"

tbl_bf <- twpi_bf |>
  mutate(year = lubridate::year(RecordDate)) |>
  group_by(Month, Year) |>
  count(Marks) |>
  mutate(Month = lubridate::month(as.numeric(Month), label = TRUE),
         `%` = round(n / sum(n) * 100, 2)) |>
  pivot_wider(names_from = Marks, values_from = c(n, `%`)) |>
  select(Year, Month, n_1, `%_1`, n_2, `%_2`, n_3, `%_3`) |>
  arrange(Year) |>
  rename(Within = n_1,
         `% Within` = `%_1`,
         At = n_2,
         `% At` = `%_2`,
         Outside = n_3,
         `% Outside` = `%_3`) |>
  ungroup()

tbl_bf |>
  mutate(Within = format(Within, big.mark = ","),
         At = format(At, big.mark = ","),
         Outside = format(Outside, big.mark = ",")) |>
  kbl(align = "r") |>
  kable_styling(full_width = FALSE)
```

Looking at the proportion of `at-the-spread` crossings after the rule change, we can see from @tbl-af that the percentages constantly low under 2%, except the first two months, June and July. (May 2013 here only accounts for 5 days). If we look at the proportion of total `at-the-spread` crossings from 26 May 2013 to December 2015, the percentage was only 1.23% (see @tbl-bfaf).

It is also worth mentioned that the monthly frequency of `within-the-spread` crossings in this period were smaller than the previous period. The frequency were mostly under 150,000 a month, except for the first five months. While the previous period account for a over 150,000 most of the time. 

On the other hand, the percentages of `outside-the-spread` crossings seemed to rise month after month. The proportion were constant at 1%-2% in the year 2013 and 2014, but then they increased significantly to 3% - 8% monthly on the year 2015. 

These number and proportion in this period shows that the amendment of the rule worked as it should be, as the `at-the-spread` trades gone from the dark pool. However, there were still not 100% effective, due to the proportion of `outside-the-spread` rose to 8% at the end of the year 2015.  

```{r}
#| label: tbl-af
#| tbl-cap: "The Number and Proportion of Internal Crossings After 26 May 2013"

tbl_af <- twpi_af |>
  mutate(year = lubridate::year(RecordDate)) |>
  group_by(Month, Year) |>
  count(Marks) |>
  mutate(Month = lubridate::month(as.numeric(Month), label = TRUE),
         `%` = round(n / sum(n) * 100, 2)) |>
  pivot_wider(names_from = Marks, values_from = c(n, `%`)) |>
  select(Year, Month, n_1, `%_1`, n_2, `%_2`, n_3, `%_3`) |>
  arrange(Year) |>
  rename(Within = n_1,
         `% Within` = `%_1`,
         At = n_2,
         `% At` = `%_2`,
         Outside = n_3,
         `% Outside` = `%_3`) |>
  ungroup()

tbl_af |>  
  mutate(Within = format(Within, big.mark = ","),
         At = format(At, big.mark = ","),
         Outside = format(Outside, big.mark = ",")) |>
  kbl(align = "r") |>
  kable_styling(full_width = FALSE)
  

```

```{r}
#| label: tbl-bfaf
#| tbl-cap: "The Percentage of Within/At/Outside-the-Spread Internal Crossings Before and After the Rule Change"

twpi_bf |>
  count(Marks) |>
  mutate(Count = n,
         Percentage = round(Count / sum(Count) * 100, 2)) |>
  select(Marks, Count, Percentage) |>
  left_join(twpi_af |>
    count(Marks) |>
    mutate(Count = n,
           Percentage = round(Count / sum(Count) * 100, 2)) |>
    select(Marks, Count, Percentage), by = "Marks") |>
  select(Marks, Percentage.x, Percentage.y) |>
  rename(Prop_Before = Percentage.x,
         Prop_After = Percentage.y) |>
  mutate(Marks = c("Within", "At Spread", "Outside")) |>
  kbl(align = "lrrrr") |>
  kable_styling(full_width = FALSE)
```

@fig-prop exhibits the movement better. As we can see from the barplot that the proportion of `at-the-spread` crossings significantly decrease after May 2013. The proportion then dominated by `within-the-spread` crossings, as the rule amendment only allow meaningful price improvement on the dark pool. There were still some breach on the TWPI rule, and the proportion were increasing towards the end of 2015, however the number and proportion were not significant compared to the compliance dark trades.

```{r}
#| label: fig-prop
#| fig-cap: "Proportion of each trade price groups from Januari 2012 until December 2015. Red dash line is when the meaningful price improvement rule introduced."

combined <- rbind(tbl_bf, tbl_af) |>
  group_by(Year, Month) |>
  summarise(Within = sum(Within),
            At = sum(At),
            Outside = sum(Outside),
            `% Within` = round(Within / sum(Within, At, Outside) * 100, 2),
            `% At` = round(At / sum(Within, At, Outside) * 100, 2),
            `% Outside` = round(Outside / sum(Within, At, Outside) * 100, 2)) |>
  mutate(Date = lubridate::ym(paste(Year, Month))) |>
  ungroup() |>
  select(6:9) |>
  pivot_longer(cols = 1:3, names_to = "Price", values_to = "Proportion") 
  
combined |> 
  mutate(Price = ifelse(Price == "% Within", "Within-the-spread", ifelse(Price == "% At", "At-the-spread", "Outside-the-spread")),
         Price = factor(Price, levels = c("Within-the-spread", "At-the-spread", "Outside-the-spread"))) |>
  ggplot() +
  geom_col(aes(x = Date, y = Proportion, fill = Price)) +
  geom_vline(xintercept = as.Date("2013-05-01"), color = "blue", linetype = "solid", size = 1) +
  annotate("text", x = as.Date("2013-05-01"), y = 75, label = "TWPI Rule", color = "blue", hjust = -0.1, size = 4) +
  theme_bw() 

```

*Economic Implication*

```{r, eval = FALSE}
twpi5 <- twpi_df |>
  select(2,4,17, 25) |>
  mutate(RecordDate = lubridate::ymd(RecordDate))

twpi5 <- twpi5 |> # remove duplicates
  group_by(Mykey) |>
  filter(n() == 1) |>
  ungroup()

save(twpi5, file = "data/report/twpi5")

```

```{r, eval = FALSE}
twpi5 |> 
  filter(RecordDate > (lubridate::ymd(20121231)) &
           RecordDate < lubridate::ymd(20130526)) |>
  group_by(Marks) |>
  summarise(Dollar = sum(DollarValue))
  
twpi5 |> 
  filter(RecordDate < (lubridate::ymd(20131014)) &
           RecordDate > lubridate::ymd(20130526)) |>
  group_by(Marks) |>
  summarise(Dollar = sum(DollarValue))
```

Comparing the dollar value of the trades can be useful to examine the economic implication of the rule change. We compare 100 days before and after the rule change in 26 May 2013. The reason of using smaller period of time is to control the change in trend of transaction, as well as minimize the the impact of macro and microeconomics change. One hundred trading days before TWPI rule amendment, `at-the-spread` crossings has value of \$3.97 billion, while 100 trading days after the value dropped to just \$216 million. This decrease in value indicates that there was a movement of around almost \$3.7 billion from `dark pool` to `lit pool`, from the stocks listed on ASX200 index alone.

This number was quite high, given that this was only from the stocks listed on ASX200 index, and only account for 100 trading days. With this number alone, ASX could charge around \$45,000 based on the trading fee on `lit market` around 0.15 basis point. The number will rise much higher if we consider the whole transaction on all stocks listed on ASX.

From the report by Charles Lane Advisory, *Review of recent rule changes affecting dark liquidity* (2014), it is known that there was a decline of total dollar volume in the below block size dark pool 100 days before and after the rule change, from \$54.9 billion to \$21 billion, meaning that there was a movement of more than \$33 billion from `dark pool` to `lit pool`. However, these numbers have taken into account all stocks in ASX and the priority crossings mechanism occurred on the ASX trading platform, in which we will not consider in this study. Based on these numbers, the `at-the-spread` crossings occurred on the ASX200 index listed companies only account for approximately 12% of that from all stocks listed on ASX.

## Violation Pattern After the Rule Change

- there still violation, only around 5%, show barplot monthly
- any broker dominates (show top 10)? show barplot, anonym the broker id
- 2 dominates the TWPI, how much? tell that no media release of penalty to twpi rule violation in this period
- what kind of broker who violate the rule? since there are 2 dominating the violation, the broker type may have nothing to do with the tendency. it depends on each broker apparently 

*Modelling*

- we use data from the 2 dominated brokers to train the model, because if not, the pattern will tend to find the broker id as dominant variable causing violation. split the data to train and test
- then check robustness to other brokers data
- remind the imbalances in the data

*Logistic Regresion*

- explain how to find the best threshold
- show the confusion matrix of the test set
- show the table for the balance accuracy, sensitivity and specificity of the test set
- show the significance of the variables
- explain the contribution of the significant variables

*Random Forest*

- explain step downsample, because of the imbalance in the data
- show the confusion matrix of the test set
- show the table for the balance accuracy, sensitivity and specificity of the test set
- show the variable importance, explain

*Robustness Check*

- show the confusion matrix of both model
- show the balance accuracy, sensitivity and specificity
- explain which model is preferred





## Analysis

```{r, eval = FALSE}
# sort top 10 broker 

broker_10 <- twpi_af |> 
  count(BuyerBrokerID) |>
  arrange(desc(n)) |>
  head(10) |>
  pull(BuyerBrokerID)

twpi_af |> 
  filter(BuyerBrokerID %in% broker_10) |>
  group_by(Marks) |>
  count(BuyerBrokerID) |>
  mutate(percentage = round(n / sum(n) * 100), 2) |>
  ungroup() |>
  filter(Marks != 1) |>
  mutate(Marks = ifelse(Marks == 2, "At Spread", "Outside")) |>
  ggplot(aes(x = reorder(BuyerBrokerID, -n), y = percentage, fill = Marks)) +
  geom_col(position = "dodge") +
  labs(title = "Violation in Percentage to TWPI",
       x = "Broker ID",
       y = "Percentage to TWPI") +
  theme_bw()
```

```{r}
# prepare data to model, only top 2 brokers

# data <- twpi_af |> 
#   filter(BuyerBrokerID %in% c(210,361),
#          Hour != 16) |>
#   # left_join(broker2[, c("No", "Description")], by = c("BuyerBrokerID" = "No")) |>
#   mutate(Violation = as.factor(ifelse(Marks == 1, 0, 1)),
#          Violation_1 = as.factor(ifelse(Marks == 2, 1, 0)),
#          Violation_2 = as.factor(ifelse(Marks == 3, 1, 0))) |>
#   select(2, 6, 8:16, 18:20) |>
#   na.omit()

# save(data, file = "data/report/data")

load("data/report/data")

viol_at <- data |> select(-12, -14) 
viol_out <- data |> select(-12, -13) 

set.seed(1010)
va_split <- initial_split(viol_at, 2/3, strata = Violation_1)
va_train <- training(va_split)
va_test <- testing(va_split)

vo_split <- initial_split(viol_out, 2/3, strata = Violation_2)
vo_train <- training(vo_split)
vo_test <- testing(vo_split)

```

### Logistic Regression

```{r, eval = FALSE}
log_mod <- logistic_reg() |>
  set_engine("glm") |>
  set_mode("classification") |>
  translate()
log_fit1 <- log_mod |>
  fit(Violation_1 ~ .,
      data = va_train[, -c(1, 8, 9)])
log_fit2 <- log_mod |>
  fit(Violation_2 ~ .,
      data = vo_train[, -c(1, 8, 9)])

tidy(log_fit1) 
glance(log_fit1)

tidy(log_fit2) 
glance(log_fit2)

saveRDS(log_fit1, file = "data/report/log_fit1.rds")
saveRDS(log_fit2, file = "data/report/log_fit2.rds")
```

```{r}
# measure accuracy logistic regression on at-spread violation 
log_fit1 <- readRDS("data/report/log_fit1.rds")


# choose threshold
log1_pred <- log_fit1 |> 
  augment(new_data = va_train) |>
  select(1:3, 15)

roc_log1 <- roc(log1_pred$Violation_1, log1_pred$.pred_1)

# plot(roc_log1, col = "blue", main = "ROC Curves")

roc_log1_optimal_threshold <- coords(roc_log1, "best", ret="threshold")


```

```{r}
va_tr_pred <- log_fit1 |> 
  augment(new_data = va_train) |>
  mutate(pred_viol = as.factor(ifelse(.pred_1 > roc_log1_optimal_threshold[1,1], 1, 0)))
va_ts_pred <- log_fit1 |> 
  augment(new_data = va_test) |>
  mutate(pred_viol = as.factor(ifelse(.pred_1 > roc_log1_optimal_threshold[1,1], 1, 0)))

matrix_va_tr <- va_tr_pred |>
  count(Violation_1, pred_viol) |>
  group_by(Violation_1) |>
  mutate(cl_acc = n[pred_viol == Violation_1]/sum(n)) |>
  pivot_wider(names_from = pred_viol, 
              values_from = n, values_fill=0)

matrix_va_ts <- va_ts_pred |>
  count(Violation_1, pred_viol) |>
  group_by(Violation_1) |>
  mutate(cl_acc = n[pred_viol == Violation_1]/sum(n)) |>
  pivot_wider(names_from = pred_viol, 
              values_from = n, values_fill=0)

bal_accuracy_logistic_va <- bal_accuracy(va_ts_pred, Violation_1, pred_viol)$.estimate

```

Confusion Matrix and Statistics of `at-the-spread` violations on the test dataset:

```{r}
# confusion matrix best threshold
confusionMatrix(va_ts_pred$pred_viol, va_ts_pred$Violation_1)
```

```{r}
# measure accuracy logistic regression on outside violation 
log_fit2 <- readRDS("data/report/log_fit2.rds")

# choose threshold
log2_pred <- log_fit2 |> 
  augment(new_data = vo_train) |>
  select(1:3, 15)

roc_log2 <- roc(log2_pred$Violation_2, log2_pred$.pred_1)

# plot(roc_log2, col = "blue", main = "ROC Curves")

roc_log2_optimal_threshold <- coords(roc_log2, "best", ret="threshold")
```


```{r}
vo_tr_pred <- log_fit2 |> 
  augment(new_data = vo_train) |>
  mutate(pred_viol = as.factor(ifelse(.pred_1 > roc_log2_optimal_threshold[1,1], 1, 0)))
vo_ts_pred <- log_fit2 |> 
  augment(new_data = vo_test) |>
  mutate(pred_viol = as.factor(ifelse(.pred_1 > roc_log2_optimal_threshold[1,1], 1, 0)))

matrix_vo_tr <- vo_tr_pred |>
  count(Violation_2, pred_viol) |>
  group_by(Violation_2) |>
  mutate(cl_acc = n[pred_viol == Violation_2]/sum(n)) |>
  pivot_wider(names_from = pred_viol, 
              values_from = n, values_fill=0)

matrix_vo_ts <- vo_ts_pred |>
  count(Violation_2, pred_viol) |>
  group_by(Violation_2) |>
  mutate(cl_acc = n[pred_viol == Violation_2]/sum(n)) |>
  pivot_wider(names_from = pred_viol, 
              values_from = n, values_fill=0)


bal_accuracy_logistic_vo <- bal_accuracy(vo_ts_pred, Violation_2, pred_viol)$.estimate

```

Confusion Matrix and Statistics of `outside-spread` violations on the test dataset:

```{r}
# confusion matrix best threshold
confusionMatrix(vo_ts_pred$pred_viol, vo_ts_pred$Violation_2)
```

#### Apply model to twpi data outside broker 210 and 361

```{r}
# filter data
# data_2 <- twpi_af |> 
#   filter(!(BuyerBrokerID %in% c(210,361)),
#          Hour != 16) |>
#   # left_join(broker2[, c("No", "Description")], by = c("BuyerBrokerID" = "No")) |>
#   mutate(Violation = as.factor(ifelse(Marks == 1, 0, 1)),
#          Violation_1 = as.factor(ifelse(Marks == 2, 1, 0)),
#          Violation_2 = as.factor(ifelse(Marks == 3, 1, 0))) |>
#   select(2, 6, 8:16, 18:20) |>
#   na.omit()
# 
# save(data_2, file = "data/report/data_2")

load("data/report/data_2")
```

```{r}
# apply at-spread violation model and measure accuracy 

data_2_va_pred <- log_fit1 |> 
  augment(new_data = data_2) |>
  mutate(pred_viol = as.factor(ifelse(.pred_1 > roc_log1_optimal_threshold[1,1], 1, 0)))

matrix_data_2_va <- data_2_va_pred |>
  count(Violation_1, pred_viol) |>
  group_by(Violation_1) |>
  mutate(cl_acc = n[pred_viol == Violation_1]/sum(n)) |>
  pivot_wider(names_from = pred_viol, 
              values_from = n, values_fill=0)


bal_accuracy_logistic_data2_1 <- bal_accuracy(data_2_va_pred, Violation_1, pred_viol)$.estimate
bal_accuracy_logistic_data2_1
```

Confusion Matrix and Statistics of `at-the-spread` violations on the new dataset:

```{r}
# confusion matrix best threshold
confusionMatrix(data_2_va_pred$pred_viol, data_2_va_pred$Violation_1)
```

```{r}
# apply outside violation model and measure accuracy 

data_2_vo_pred <- log_fit2 |> 
  augment(new_data = data_2) |>
  mutate(pred_viol = as.factor(ifelse(.pred_1 > roc_log2_optimal_threshold[1,1], 1, 0)))

matrix_data_2_vo <- data_2_vo_pred |>
  count(Violation_2, pred_viol) |>
  group_by(Violation_2) |>
  mutate(cl_acc = n[pred_viol == Violation_2]/sum(n)) |>
  pivot_wider(names_from = pred_viol, 
              values_from = n, values_fill=0)


bal_accuracy_logistic_data2_2 <- bal_accuracy(data_2_vo_pred, Violation_2, pred_viol)$.estimate
bal_accuracy_logistic_data2_2
```

Confusion Matrix and Statistics of `outside-spread` violations on the new dataset:

```{r}
# confusion matrix best threshold
confusionMatrix(data_2_vo_pred$pred_viol, data_2_vo_pred$Violation_2)
```

### Random Forest

#### At-The-Spread

```{r}
data_rec1 <- recipe(Violation_1 ~ ., data = va_train) |>
  step_downsample(Violation_1) |>
  step_rm(Instrument, Year, Month)
```

```{r}
# data_prep1 <- prep(data_rec1)
# saveRDS(data_prep1, file = "data/report/data_prep1.rds")

data_prep1 <- readRDS("data/report/data_prep1.rds")
data_juice1 <- juice(data_prep1)
```

```{r, eval = FALSE}
# tuning hyperparameters
rf_tune1 <- rand_forest(
  mtry = tune(),
  trees = 1000,
  min_n = tune()) |>
  set_mode("classification") |>
  set_engine("randomForest")

tune_wf1 <- workflow() |>
  add_recipe(data_rec1) |>
  add_model(rf_tune1)
  
doParallel::registerDoParallel() # for parallel processing

set.seed(3344)
w_folds1 <- vfold_cv(va_train)

rf_grid1 <- grid_regular(mtry(range = c(2, 8)),
                        min_n(range = c(1, 4)),
                        levels = 3)

tune_res1 <- 
  tune_grid(tune_wf1,
            resamples = w_folds1,
            grid = rf_grid1
            )

# select best combination based on accuracy
best_hp1 <- tune_res1 |> select_best(metric = "accuracy")
```

```{r, eval = FALSE}
# fit the model
set.seed(3344)
rf1 <- rand_forest(mtry = 5, 
                     min_n = 4) |>
  set_mode("classification") |>
  set_engine("randomForest", importance = TRUE)

rf_workflow1 <- workflow() |>
  add_model(rf1) |>
  add_formula(Violation_1 ~ .)

fit_rf1 <- rf_workflow1 |> 
  fit(data = data_juice1)

saveRDS(fit_rf1, file = "data/report/fit_rf1.rds")
```

```{r}
# apply the model to test set
fit_rf1 <- readRDS("data/report/fit_rf1.rds")

va_ts_pred_rf <- bake(data_prep1, va_test) |>
  mutate(pviol = predict(fit_rf1,
                         bake(data_prep1, va_test))$.pred_class)
```

Variable importance of `at-the-spread` violations:

```{r}
# measure accuracy

cf_rf_ts1 <- va_ts_pred_rf |>
  count(Violation_1, pviol) |>
  group_by(Violation_1) |>
  mutate(Accuracy = n[Violation_1==pviol]/sum(n)) |>
  pivot_wider(names_from = "pviol", 
              values_from = n, values_fill = 0)


bal_accuracy_rf1 <- bal_accuracy(va_ts_pred_rf, Violation_1, pviol)$.estimate

rf_model1 <- pull_workflow_fit(fit_rf1)$fit
importance(rf_model1)
```

Confusion Matrix and Statistics of `at-the-spread` violations on the test dataset:

```{r}
confusionMatrix(va_ts_pred_rf$pviol, va_ts_pred_rf$Violation_1)
```

#### Outside the Spread

```{r}
data_rec2 <- recipe(Violation_2 ~ ., data = vo_train) |>
  step_downsample(Violation_2) |>
  step_rm(Instrument, Year, Month)
```

```{r}
# data_prep2 <- prep(data_rec2)
# saveRDS(data_prep2, file = "data/report/data_prep2.rds")

data_prep2 <- readRDS("data/report/data_prep2.rds")
data_juice2 <- juice(data_prep2)
```

```{r, eval = FALSE}
# tuning hyperparameters
rf_tune2 <- rand_forest(
  mtry = tune(),
  trees = 1000,
  min_n = tune()) |>
  set_mode("classification") |>
  set_engine("randomForest")

tune_wf2 <- workflow() |>
  add_recipe(data_rec2) |>
  add_model(rf_tune2)
  
doParallel::registerDoParallel() # for parallel processing

set.seed(3344)
w_folds2 <- vfold_cv(vo_train)

rf_grid2 <- grid_regular(mtry(range = c(2, 8)),
                        min_n(range = c(1, 4)),
                        levels = 3)

tune_res2 <- 
  tune_grid(tune_wf2,
            resamples = w_folds2,
            grid = rf_grid2
            )

# select best combination based on accuracy
best_hp2 <- tune_res2 |> select_best(metric = "accuracy")
```

```{r, eval = FALSE}
# fit the model
set.seed(3344)
rf2 <- rand_forest(mtry = 5, 
                     min_n = 4) |>
  set_mode("classification") |>
  set_engine("randomForest", importance = TRUE)

rf_workflow2 <- workflow() |>
  add_model(rf2) |>
  add_formula(Violation_2 ~ .)

fit_rf2 <- rf_workflow2 |> 
  fit(data = data_juice2)

saveRDS(fit_rf2, file = "data/report/fit_rf2.rds")
```

```{r}
# apply the model to test set
fit_rf2 <- readRDS("data/report/fit_rf2.rds")

vo_ts_pred_rf <- bake(data_prep2, vo_test) |>
  mutate(pviol = predict(fit_rf2,
                         bake(data_prep2, vo_test))$.pred_class)
```

Variable importance of `outside-the-spread` violation:

```{r}
# measure accuracy
cf_rf_ts2 <- vo_ts_pred_rf |>
  count(Violation_2, pviol) |>
  group_by(Violation_2) |>
  mutate(Accuracy = n[Violation_2==pviol]/sum(n)) |>
  pivot_wider(names_from = "pviol", 
              values_from = n, values_fill = 0)

bal_accuracy_rf2 <- bal_accuracy(vo_ts_pred_rf, Violation_2, pviol)$.estimate

rf_model2 <- pull_workflow_fit(fit_rf2)$fit
importance(rf_model2)
```

Confusion Matrix and Statistics of `outside-spread` violations on the test dataset:

```{r}
confusionMatrix(vo_ts_pred_rf$pviol, vo_ts_pred_rf$Violation_2)
```

#### Apply model to twpi data outside broker 210 and 361

```{r}
# apply at-spread violation model and measure accuracy 
data_2_va_pred_rf <- bake(data_prep1, data_2) |>
  mutate(pviol = predict(fit_rf1,
                         bake(data_prep1, data_2))$.pred_class)

cf_rf_data_2_1 <- data_2_va_pred_rf |>
  count(Violation_1, pviol) |>
  group_by(Violation_1) |>
  mutate(Accuracy = n[Violation_1==pviol]/sum(n)) |>
  pivot_wider(names_from = "pviol", 
              values_from = n, values_fill = 0)

bal_accuracy_rf_data2_1 <- bal_accuracy(data_2_va_pred_rf, Violation_1, pviol)$.estimate

```

Confusion Matrix and Statistics of `at-the-spread` violations on the new dataset:

```{r}
confusionMatrix(data_2_va_pred_rf$pviol, data_2_va_pred_rf$Violation_1)
```

```{r}
# apply outside violation model and measure accuracy 
data_2_vo_pred_rf <- bake(data_prep2, data_2) |>
  mutate(pviol = predict(fit_rf2,
                         bake(data_prep2, data_2))$.pred_class)

cf_rf_data_2_2 <- data_2_vo_pred_rf |>
  count(Violation_2, pviol) |>
  group_by(Violation_2) |>
  mutate(Accuracy = n[Violation_2==pviol]/sum(n)) |>
  pivot_wider(names_from = "pviol", 
              values_from = n, values_fill = 0)

bal_accuracy_rf_data2_2 <- bal_accuracy(data_2_vo_pred_rf, Violation_2, pviol)$.estimate

```

Confusion Matrix and Statistics of `outside-spread` violations on the new dataset:

```{r}
confusionMatrix(data_2_vo_pred_rf$pviol, data_2_vo_pred_rf$Violation_2)
```

### Summary

```{r}

summary <- data.frame(
  Violation = c("at-spread", "outside-spread"),
  GLM_top2 = c(bal_accuracy_logistic_va, bal_accuracy_logistic_vo),
  GLM_others = c(bal_accuracy_logistic_data2_1, bal_accuracy_logistic_data2_2),
  RF_top2 = c(bal_accuracy_rf1, bal_accuracy_rf2),
  RF_others = c(bal_accuracy_rf_data2_1, bal_accuracy_rf_data2_2)
)


summary |>
  kbl() |>
  kable_styling(full_width = FALSE, position = "left")
```

### Limitation

# Conclusion 500

## Potential Future Work

# References

ASIC Market Integrity Rules (Competition in Exchange Markets) 2011.

Australian Securities and Investments Commission. (20 May 2004). *Market demutualisation and privatisation: The Australian experience* [Speech]. Retrieved 24 October 2024, from https://asic.gov.au/about-asic/news-centre/speeches/market-demutualisation-and-privatisation-the-australian-experience/

Australian Securities and Investments Commission. (May 2014). *Report 394: Review of recent rule changes affecting dark liquidity* (PDF). Retrieved 24 October 2024, from https://download.asic.gov.au/media/1344596/rep394-published-19-May-2014.pdf

Australian Securities and Investments Commission. (May 2022). *Market integrity update: Issue 137*. Retrieved 24 October 2024, from https://asic.gov.au/about-asic/corporate-publications/newsletters/market-integrity-update/miu-issue-137-may-2022/

Australian Securities and Investments Commission. (3 February 2023). *Wilson's Advisory and Stockbroking Ltd ACN 010 529 665 pays $548,328 infringement notice* [Media release]. Retrieved 24 October 2024, from https://asic.gov.au/about-asic/news-centre/find-a-media-release/2023-releases/23-016mr-wilsons-advisory-and-stockbroking-ltd-acn-010-529-665-pays-548-328-infringement-notice/

Australian Securities Exchange. (n.d.). ASX regulatory framework. Retrieved 24 October 2024, from https://www.asx.com.au/about/regulation

Australian Securities Exchange. (n.d.). ASX story. Retrieved 24 October 2024, from https://www.asx.com.au/about/asx-story

Australian Securities Exchange. (n.d.). About Us. Retrieved 24 October 2024, from https://www.asx.com.au/about#:~:text=As%20an%20integrated%20exchange%2C%20ASX's,equities%20and%20fixed%20income%20markets.

Australian Securities Exchange. (19 February 2024). *ASX Trade: Markets Participant and Trading Schedule of Fees* (PDF). Retrieved 24 October 2024, from https://asxonline.com/content/dam/asxonline/public/documents/schedule-of-fees/asx-trade-markets-participant-and-trading-schedule-of-fees.pdf

Charles Lane Advisory Pty Ltd. (May 2014). *Report by Charles Lane Advisory: Review of recent rule changes affecting dark liquidity* (PDF). Attachment to ASIC Report 394: Review of recent rule changes affecting dark liquidity. Retrieved 24 October 2024, from https://download.asic.gov.au/media/1344590/rep394-attachment-published-19-May-2014.pdf

Duong, H. N., Lajbcygier, P., Lu, J. S., & Vu, V. H. (2018). The effect of anonymity on price efficiency: Evidence from the removal of broker identities. *Pacific-Basin Finance Journal, 51*, 95–107. https://doi.org/10.1016/j.pacfin.2018.06.004

Securities Industry Research Centre of Asia-Pacific. (n.d.). *About SIRCA*. SIRCA. Retrieved 26 October 2024, from https://www.sirca.org.au/about-sirca/
