---
title: "Trade With Price Improvement"
author: "Bagas Trilaksonoaji"
date: "`r Sys.Date()`"
format: 
  pdf:
    toc: true
    number-sections: true
execute: 
  eval: TRUE
  echo: false
  message: false
  warning: false
cache: TRUE
editor_options: 
  chunk_output_type: inline
---

\pagebreak

# Abstract

# Introduction 

There were numerous equities exchanges established in the long history of Australia. Almost every big city in Australia once had their own stock exchange, namely Sydney, Brisbane, Melbourne, Adelaide, Perth, and even Hobart. The Australian Parliament merged all six stock exchanges in 1987, creating the Australian Stock Exchange [(Australian Stock Exchange, n.d.)](https://www.asx.com.au/about/asx-story). Two decades later, this equities exchange merged with Sydney Futures Exchange to become the Australian Securities Exchange. Although its official name is now ASX Limited, it is more convenient to simply refer to it as ASX. In this report, however, we will use the term ASX to refer to the company or the equities market.

As a market operator, ASX's main tasks include reviewing listing applications, supervising listed entities as well as listed brokers, operating and maintaining an electronic trading platform for buying and selling stocks, and providing clearing and settlement services. [(Australian Stock Exchange, n.d.)](https://www.asx.com.au/about#:~:text=As%20an%20integrated%20exchange%2C%20ASX's,equities%20and%20fixed%20income%20markets.).

ASX operates under the regulatory framework of two government agencies, the Australia Securities and Investments Commission (ASIC) and the Reserve Bank of Australia (RBA) [(Australian Stock Exchange, n.d.)](https://www.asx.com.au/about/regulation). ASIC oversees stock market trading and enforces financial market conduct laws, while the Reserve Bank of Australia (RBA) oversees clearing and settlement-related services.

## Background and Motivation 

Anyone who wants to buy or sell stocks can place an order on the ASX trading platform. When the price of the buy and sell side matches, the trade can be executed. Outstanding orders will remain on the platform and visible to market participants. This mechanism promotes transparency and market equilibrium in supply and demand, thereby ensuring fair stock price formation [(Australian Securities and Investments Commission, 2022)](https://asic.gov.au/about-asic/corporate-publications/newsletters/market-integrity-update/miu-issue-137-may-2022/). That is why we often call this trading venue provided by the ASX trading platform `lit market` or `lit pool`.

On the other hand, there are other venues that do not disclose the orders placed, at least until the transaction is executed. We will use the term `dark pool` to refer to this kind of trading venue. There are several `dark pools` available for stock trading in Australia. However, in this project, we will focus on the `dark pool` executed on the broker's internal crossing system. This kind of `dark pool` is often called NBBO trade, which stands for National Best Bid and Offer Prices. 

Crossing system is an electronic platform provided by the brokers to match orders placed by their clients with their other clients or with their own account [(Australian Securities and Investments Commission, 2014)](https://download.asic.gov.au/media/1344596/rep394-published-19-May-2014.pdf). In other words, orders can be matched on the broker's internal system before they are passed through to the ASX trading platform. For the next section, we will use the term `crossing` to refer to the trades in the broker's internal crossing system.

What is the benefit for brokers and traders if they trade on the crossing system instead of the ASX `lit pool`? One of the primary benefits is the significantly lower trading fees. Trades on the ASX platform are charged about 0.15 basis points per trade, whereas internal crossing charges as much as 0.04 basis points per trade, capped at $1,000 per month [(Australian Securities Exchange, 2024)](https://asxonline.com/content/dam/asxonline/public/documents/schedule-of-fees/asx-trade-markets-participant-and-trading-schedule-of-fees.pdf)). The difference is quite huge given how much a broker can trade on a daily basis.

It is important to note that the rule in which internal crossings referred to was differed before and after 26 May 2013:

1. **Before**

    Rule 4.2.3(1)(b) of the ASIC Market Integrity Rules (Competition in Exchange Markets) 2011, regulates the `At or Within the Spread` trades. As the name suggests, the executed trade price must be at the same as the best available bid or best available offer of the `lit market` (at-the-spread), or better (within-the-spread). This implies that the price executed in this "dark pool" cannot be lower than the best bid price or higher than the best offer price in the "lit market".
2. **After**

    The rule had changed, based on Rule 4.2.3 of the ASIC Market Integrity Rules (Competition in Exchange Markets) Amendment 2012 (No. 1), the trades on the crossing system must provide meaningful price improvement [(Australian Securities and Investments Commission, 2014)](https://download.asic.gov.au/media/1344596/rep394-published-19-May-2014.pdf). Trades at the best bid or best offer price of the `lit market` are not in the option anymore. This rule was effective on 26 May 2013. ASIC and ASX refer to this rule as `Trades With Price Improvement`, or next we can refer to this as TWPI. This study will specifically focus on the internal crossings, which follow the TWPI rule.

The amendment of the rule in May 2013 was to address the negative impact of the dark trades on the price formation, specifically from the `at-the-spread` trades [(Australian Securities and Investments Commission, 2014)](https://download.asic.gov.au/media/1344596/rep394-published-19-May-2014.pdf), due to traders may prefer to do the trades on the `dark pool` as it costs less fees compared to the `lit pool`. The adverse effect on price discovery was one of the main reasons why the regulator wanted more trading to occur in the `lit pool`.

On the other hand, the ASX was also motivated by an economic incentive. On October 14, 1998, ASX became a publicly owned company after demutualising itself. The main analysis behind this was that "*ASX needed to become more flexible, responsive, and commercially focused, capable of quickly taking up emerging commercial opportunities*" [(Australian Securities and Investments Commission, 2004)](https://asic.gov.au/about-asic/news-centre/speeches/market-demutualisation-and-privatisation-the-australian-experience/). Therefore, ASX, being an all-for-profit enterprise, would seize any business opportunity to increase its revenue and maximize its returns to shareholders. ASIC amended the rules to provide meaningful price improvements, which would shift a significant portion of dark trades to the ASX platform, thereby increasing its profit from trading fees.

**Key Motivation**

In May 2022, ASIC published a Market Integrity Update (Issue 137), which included a review of TWPI. In the article, ASIC mentioned that several participants were reporting a significant number of their TWPIs without any price improvement. ASIC emphasized that such trades must be done in the `lit market` because they will promote liquidity and price formation. TWPI is an exception because it offers better prices for the investor—a price that the `lit market` cannot provide.

Following up on the Market Integrity Update, ASIC issued an infringement notice to Wilson Advisory and Stockbroking Pty Ltd, requiring them to pay $548,328 to ASIC on behalf of the Commonwealth. The notice was given on 16 December 2022, while published on an ASIC media release on 3 February 2023.

The penalty shows that ASIC has started to take the breach of the TWPI rule more seriously. Probably due to the increasing number of dark trades that should be done on the `lit market`, as ASIC intention of rule change in 2013 was to increasing the proportion of trading on the `lit market`, hence, enhancing the fairness on the market price formation [Australian Securities and Investments Commission, 2014](https://download.asic.gov.au/media/1344596/rep394-published-19-May-2014.pdf)

Based on these insights, I'm intrigued to analyse the `at-the-spread` dark trades before and after the rule amendment in 26 May 2013. Different from the ASIC report ([Report 394](https://download.asic.gov.au/media/1344596/rep394-published-19-May-2014.pdf)) on May 2014, this study will focus on the internal crossings mechanism, among all other dark pool mechanisms. Furthermore, the internal crossings violation on the TWPI rule will be analysed, along with the model that explains the pattern of the violation.

## Purpose and Key Questions

The main objective of this study is to look into how the rule change in 2013 affected `at-the-spread` crossings reported as TWPI. The way brokers act and trade on the `dark pool` should be greatly affected by this change, as they must move the `at-the-spread` crossings to the `lit pool`. Moreover, this study will also examine the violation of the TWPI after the rule change.

The study will compare transaction data before and after the 2013 regulation change, with a focus on `at-the-spread` crossings. Trade volumes, prices, time stamps, broker IDs, and several other variables will be included in the analysis. There will be simple statistical methods used to compare trade distributions, and machine learning methods may be used to find trends of the violations in the TWPI rule.

### Understand the Impact on At-the-Spread Crossings

`At-the-spread` crossings were possible before the rule change, which meant that trades could be made at the best bid or best offer price available on the `lit market`. However, the 2013 amendment says that trades must lead to price improvements, in other words, `at-the-spread` crossings are no longer allowed. The main question to ask is the proportion of `at-the-spread` crossings among all TWPI-reported trades before and after 26 May 2013. This comparison will help to figure out whether the rule change stopped people from doing `at the spread` crossings at the `dark pool`. Furthermore, the dollar value that possibly moved to the `lit market` from the `dark pool` will also be analysed.

**Key Question**:

1. Did the number of `at-the-spread` crossings drop significantly after 26 May 2013?
2. How much was approximately the value of `at-the-spread` internal crossings that moved to the lit market after the rule change?

### Identify the Violation Pattern

ASIC's media release and publication in the last couple of years suggested that there have been TWPI rule violations in recent years. This study will examine the monetary worth of the violations, identify which brokers do the most violations, and find any pattern by creating a model to help spot non-compliance in the future.

**Key Question**:

1. How many violations, both in terms of the quantity of trades and the dollar value, occurred following the rule change?
2. What kind of brokers violate the rules the most? Are some brokers more likely to break TWPI rules than others?
3. What conditions contributed to the occurrence of the violations?

## Significance and Contributions 500

There are three main areas in which this study may contribute: academic, regulatory, and economic.

### Economic Significance

The purpose of the rule amendment in 2013 was to encourage meaningful price improvement on the `dark pool` ([ASIC, 2014](https://download.asic.gov.au/media/1344596/rep394-published-19-May-2014.pdf)). Without price improvement, trading must be done on the ASX trading platform to promote liquidity on the market. Charles Lane Advisory Pty Ltd report ([Attachment to ASIC Report 394](https://download.asic.gov.au/media/1344590/rep394-attachment-published-19-May-2014.pdf)) in May 2014 has thoroughly analyse the impact of this rule change and stated that there was a decline in the dollar value from \$54.9 billion to \$21 billion, 100 days prior the rule change and after. Our study aims to validate this figure by focusing only on internal crossings and extending the analysis over a one-year period, with the goal of providing a more in-depth understanding of the rule's economic impacts.

From a financial point of view, it is also essential to comprehend TWPI violations since they have the potential to compromise market integrity in general [(Australian Securities and Investments Commission, 2014)](https://download.asic.gov.au/media/1344596/rep394-published-19-May-2014.pdf). By figuring out how much value in dollar TWPI violations occurred, this study can show how much value was traded not in a legal manner. If these trades had taken place on the `lit market`, then there would be trading fees that should be paid to the market operator (ASX). Internal crossing trades incur a fee of approximately 0.04 basis points per trade, whereas trades on the ASX platform cost 0.15 basis points per trade [(Australian Securities and Investments Commission, 2024)](https://asxonline.com/content/dam/asxonline/public/documents/schedule-of-fees/asx-trade-markets-participant-and-trading-schedule-of-fees.pdf). This significant difference in trading fees indicates additional economic implications.

### Regulatory Significance

Understanding the magnitude of the TWPI violation since its amendment in 2013 may provide regulators with new insights to address the TWPI reporting issues at the prevention level rather than through corrective action. Furthermore, if the detection model developed in this study proves effective, regulators might utilize it as part of an automated enforcement mechanism to ensure TWPI reporting follows the regulations.

### Academic Significance

There is a lot of academic literature on special orders and hidden orders, but this study adds a new dimension by focusing on TWPI trades. This study's novel contribution is the use of a prediction model to identify patterns and conditions that result in TWPI rule violations. The study might help academics learn more about how rules are enforced in financial markets and the situations where traders are most likely to take advantage of the system loopholes.

# Data and Methodology 1000

# Result and Discussion 2000

```{r}
library(tidyverse)
library(hms)
library(knitr)
library(kableExtra)
options(knitr.table.html.attr = "quarto-disable-processing=true")
library(tidymodels)
library(discrim)
library(rpart)
library(rpart.plot)
library(randomForest)
library(themis)
library(caret)
library(pROC)

```

```{r, eval = FALSE}
# list all csv files
list_2012 <- list.files(path = "data/2012/", pattern = "*.csv", full.names = TRUE)
list_2013 <- list.files(path = "data/2013/", pattern = "*.csv", full.names = TRUE)
list_2014 <- list.files(path = "data/2014/", pattern = "*.csv", full.names = TRUE)
list_2015 <- list.files(path = "data/2015/", pattern = "*.csv", full.names = TRUE)

# read all CSV files into a list of data frames
df_list_2012 <- lapply(list_2012, function(file) {
  read.csv(file, colClasses = c(Mykey = "character", BidID = "character", AskID = "character"))
  })
df_list_2013 <- lapply(list_2013, function(file) {
  read.csv(file, colClasses = c(Mykey = "character", BidID = "character", AskID = "character"))
  })
df_list_2014 <- lapply(list_2014, function(file) {
  read.csv(file, colClasses = c(Mykey = "character", BidID = "character", AskID = "character"))
  })
df_list_2015 <- lapply(list_2015, function(file) {
  read.csv(file, colClasses = c(Mykey = "character", BidID = "character", AskID = "character"))
  })

df_list <- c(df_list_2012, df_list_2013, df_list_2014, df_list_2015)

# combine the data frames into a single data frame
twpi_df <- bind_rows(df_list)

# save the df
save(twpi_df, file = "data/twpi_2012-2015")

```

```{r, eval = FALSE}
load("data/twpi_2012-2015")
twpi4 <- twpi_df |>
  select(-Count, -(6:11), -(14:16), -(19:21)) |>
  mutate(spread = abs(Ask_before - Bid_before),
         RecordDate = lubridate::ymd(RecordDate),
         Hour = substr(HourMinuteSecond, 1, 2),
         Year = lubridate::year(RecordDate),
         Month = lubridate::month(RecordDate),
         Date = lubridate::day(RecordDate),
         Dayofweek = lubridate::wday(RecordDate)) |>
  select(1:3, 5:11, 13:18, Marks) |>
  mutate(BuyerBrokerID = as.factor(BuyerBrokerID),
         across(12:14, as.factor),
         across(16:17, as.factor)) |>
  group_by(Instrument) |>
  mutate_if(is.numeric, function(x) (x-mean(x, na.rm = TRUE))/sd(x, na.rm = TRUE)) |>
  ungroup()

twpi4 <- twpi4 |> # remove duplicates
  group_by(Mykey) |>
  filter(n() == 1) |>
  ungroup()
```

```{r, eval = FALSE}
twpi_bf <- twpi4 |>
  filter(RecordDate < lubridate::ymd("20130526"))

twpi_af <- twpi4 |>
  filter(RecordDate >= lubridate::ymd("20130526"))
```

```{r, eval = FALSE}
# load broker classification data

broker <- read.csv("data/BrokerClassifications.csv")
class <- read.csv("data/Classifications.csv")
broker2 <- broker |> left_join(class, by = "Classification") |>
  mutate(No = as.factor(No))
```

## Analysis

```{r, eval = FALSE}
# sort top 10 broker 

broker_10 <- twpi_af |> 
  count(BuyerBrokerID) |>
  arrange(desc(n)) |>
  head(10) |>
  pull(BuyerBrokerID)

twpi_af |> 
  filter(BuyerBrokerID %in% broker_10) |>
  group_by(Marks) |>
  count(BuyerBrokerID) |>
  mutate(percentage = round(n / sum(n) * 100), 2) |>
  ungroup() |>
  filter(Marks != 1) |>
  mutate(Marks = ifelse(Marks == 2, "At Spread", "Outside")) |>
  ggplot(aes(x = reorder(BuyerBrokerID, -n), y = percentage, fill = Marks)) +
  geom_col(position = "dodge") +
  labs(title = "Violation in Percentage to TWPI",
       x = "Broker ID",
       y = "Percentage to TWPI") +
  theme_bw()
```

```{r}
# prepare data to model, only top 2 brokers

# data <- twpi_af |> 
#   filter(BuyerBrokerID %in% c(210,361),
#          Hour != 16) |>
#   # left_join(broker2[, c("No", "Description")], by = c("BuyerBrokerID" = "No")) |>
#   mutate(Violation = as.factor(ifelse(Marks == 1, 0, 1)),
#          Violation_1 = as.factor(ifelse(Marks == 2, 1, 0)),
#          Violation_2 = as.factor(ifelse(Marks == 3, 1, 0))) |>
#   select(2, 6, 8:16, 18:20) |>
#   na.omit()

# save(data, file = "data/report/data")

load("data/report/data")

viol_at <- data |> select(-12, -14) 
viol_out <- data |> select(-12, -13) 

set.seed(1010)
va_split <- initial_split(viol_at, 2/3, strata = Violation_1)
va_train <- training(va_split)
va_test <- testing(va_split)

vo_split <- initial_split(viol_out, 2/3, strata = Violation_2)
vo_train <- training(vo_split)
vo_test <- testing(vo_split)

```

### Logistic Regression

```{r, eval = FALSE}
log_mod <- logistic_reg() |>
  set_engine("glm") |>
  set_mode("classification") |>
  translate()
log_fit1 <- log_mod |>
  fit(Violation_1 ~ .,
      data = va_train[, -c(1, 8, 9)])
log_fit2 <- log_mod |>
  fit(Violation_2 ~ .,
      data = vo_train[, -c(1, 8, 9)])

tidy(log_fit1) 
glance(log_fit1)

tidy(log_fit2) 
glance(log_fit2)

saveRDS(log_fit1, file = "data/report/log_fit1.rds")
saveRDS(log_fit2, file = "data/report/log_fit2.rds")
```

```{r}
# measure accuracy logistic regression on at-spread violation 
log_fit1 <- readRDS("data/report/log_fit1.rds")


# choose threshold
log1_pred <- log_fit1 |> 
  augment(new_data = va_train) |>
  select(1:3, 15)

roc_log1 <- roc(log1_pred$Violation_1, log1_pred$.pred_1)

# plot(roc_log1, col = "blue", main = "ROC Curves")

roc_log1_optimal_threshold <- coords(roc_log1, "best", ret="threshold")


```

```{r}
va_tr_pred <- log_fit1 |> 
  augment(new_data = va_train) |>
  mutate(pred_viol = as.factor(ifelse(.pred_1 > roc_log1_optimal_threshold[1,1], 1, 0)))
va_ts_pred <- log_fit1 |> 
  augment(new_data = va_test) |>
  mutate(pred_viol = as.factor(ifelse(.pred_1 > roc_log1_optimal_threshold[1,1], 1, 0)))

matrix_va_tr <- va_tr_pred |>
  count(Violation_1, pred_viol) |>
  group_by(Violation_1) |>
  mutate(cl_acc = n[pred_viol == Violation_1]/sum(n)) |>
  pivot_wider(names_from = pred_viol, 
              values_from = n, values_fill=0)

matrix_va_ts <- va_ts_pred |>
  count(Violation_1, pred_viol) |>
  group_by(Violation_1) |>
  mutate(cl_acc = n[pred_viol == Violation_1]/sum(n)) |>
  pivot_wider(names_from = pred_viol, 
              values_from = n, values_fill=0)

bal_accuracy_logistic_va <- bal_accuracy(va_ts_pred, Violation_1, pred_viol)$.estimate

```

Confusion Matrix and Statistics of `at-the-spread` violations on the test dataset:

```{r}
# confusion matrix best threshold
confusionMatrix(va_ts_pred$pred_viol, va_ts_pred$Violation_1)
```

```{r}
# measure accuracy logistic regression on outside violation 
log_fit2 <- readRDS("data/report/log_fit2.rds")

# choose threshold
log2_pred <- log_fit2 |> 
  augment(new_data = vo_train) |>
  select(1:3, 15)

roc_log2 <- roc(log2_pred$Violation_2, log2_pred$.pred_1)

# plot(roc_log2, col = "blue", main = "ROC Curves")

roc_log2_optimal_threshold <- coords(roc_log2, "best", ret="threshold")
```


```{r}
vo_tr_pred <- log_fit2 |> 
  augment(new_data = vo_train) |>
  mutate(pred_viol = as.factor(ifelse(.pred_1 > roc_log2_optimal_threshold[1,1], 1, 0)))
vo_ts_pred <- log_fit2 |> 
  augment(new_data = vo_test) |>
  mutate(pred_viol = as.factor(ifelse(.pred_1 > roc_log2_optimal_threshold[1,1], 1, 0)))

matrix_vo_tr <- vo_tr_pred |>
  count(Violation_2, pred_viol) |>
  group_by(Violation_2) |>
  mutate(cl_acc = n[pred_viol == Violation_2]/sum(n)) |>
  pivot_wider(names_from = pred_viol, 
              values_from = n, values_fill=0)

matrix_vo_ts <- vo_ts_pred |>
  count(Violation_2, pred_viol) |>
  group_by(Violation_2) |>
  mutate(cl_acc = n[pred_viol == Violation_2]/sum(n)) |>
  pivot_wider(names_from = pred_viol, 
              values_from = n, values_fill=0)


bal_accuracy_logistic_vo <- bal_accuracy(vo_ts_pred, Violation_2, pred_viol)$.estimate

```

Confusion Matrix and Statistics of `outside-spread` violations on the test dataset:

```{r}
# confusion matrix best threshold
confusionMatrix(vo_ts_pred$pred_viol, vo_ts_pred$Violation_2)
```

#### Apply model to twpi data outside broker 210 and 361

```{r}
# filter data
# data_2 <- twpi_af |> 
#   filter(!(BuyerBrokerID %in% c(210,361)),
#          Hour != 16) |>
#   # left_join(broker2[, c("No", "Description")], by = c("BuyerBrokerID" = "No")) |>
#   mutate(Violation = as.factor(ifelse(Marks == 1, 0, 1)),
#          Violation_1 = as.factor(ifelse(Marks == 2, 1, 0)),
#          Violation_2 = as.factor(ifelse(Marks == 3, 1, 0))) |>
#   select(2, 6, 8:16, 18:20) |>
#   na.omit()
# 
# save(data_2, file = "data/report/data_2")

load("data/report/data_2")
```

```{r}
# apply at-spread violation model and measure accuracy 

data_2_va_pred <- log_fit1 |> 
  augment(new_data = data_2) |>
  mutate(pred_viol = as.factor(ifelse(.pred_1 > roc_log1_optimal_threshold[1,1], 1, 0)))

matrix_data_2_va <- data_2_va_pred |>
  count(Violation_1, pred_viol) |>
  group_by(Violation_1) |>
  mutate(cl_acc = n[pred_viol == Violation_1]/sum(n)) |>
  pivot_wider(names_from = pred_viol, 
              values_from = n, values_fill=0)


bal_accuracy_logistic_data2_1 <- bal_accuracy(data_2_va_pred, Violation_1, pred_viol)$.estimate
bal_accuracy_logistic_data2_1
```

Confusion Matrix and Statistics of `at-the-spread` violations on the new dataset:

```{r}
# confusion matrix best threshold
confusionMatrix(data_2_va_pred$pred_viol, data_2_va_pred$Violation_1)
```

```{r}
# apply outside violation model and measure accuracy 

data_2_vo_pred <- log_fit2 |> 
  augment(new_data = data_2) |>
  mutate(pred_viol = as.factor(ifelse(.pred_1 > roc_log2_optimal_threshold[1,1], 1, 0)))

matrix_data_2_vo <- data_2_vo_pred |>
  count(Violation_2, pred_viol) |>
  group_by(Violation_2) |>
  mutate(cl_acc = n[pred_viol == Violation_2]/sum(n)) |>
  pivot_wider(names_from = pred_viol, 
              values_from = n, values_fill=0)


bal_accuracy_logistic_data2_2 <- bal_accuracy(data_2_vo_pred, Violation_2, pred_viol)$.estimate
bal_accuracy_logistic_data2_2
```

Confusion Matrix and Statistics of `outside-spread` violations on the new dataset:

```{r}
# confusion matrix best threshold
confusionMatrix(data_2_vo_pred$pred_viol, data_2_vo_pred$Violation_2)
```

### Random Forest

#### At-The-Spread

```{r}
data_rec1 <- recipe(Violation_1 ~ ., data = va_train) |>
  step_downsample(Violation_1) |>
  step_rm(Instrument, Year, Month)
```

```{r}
# data_prep1 <- prep(data_rec1)
# saveRDS(data_prep1, file = "data/report/data_prep1.rds")

data_prep1 <- readRDS("data/report/data_prep1.rds")
data_juice1 <- juice(data_prep1)
```

```{r, eval = FALSE}
# tuning hyperparameters
rf_tune1 <- rand_forest(
  mtry = tune(),
  trees = 1000,
  min_n = tune()) |>
  set_mode("classification") |>
  set_engine("randomForest")

tune_wf1 <- workflow() |>
  add_recipe(data_rec1) |>
  add_model(rf_tune1)
  
doParallel::registerDoParallel() # for parallel processing

set.seed(3344)
w_folds1 <- vfold_cv(va_train)

rf_grid1 <- grid_regular(mtry(range = c(2, 8)),
                        min_n(range = c(1, 4)),
                        levels = 3)

tune_res1 <- 
  tune_grid(tune_wf1,
            resamples = w_folds1,
            grid = rf_grid1
            )

# select best combination based on accuracy
best_hp1 <- tune_res1 |> select_best(metric = "accuracy")
```

```{r, eval = FALSE}
# fit the model
set.seed(3344)
rf1 <- rand_forest(mtry = 5, 
                     min_n = 4) |>
  set_mode("classification") |>
  set_engine("randomForest", importance = TRUE)

rf_workflow1 <- workflow() |>
  add_model(rf1) |>
  add_formula(Violation_1 ~ .)

fit_rf1 <- rf_workflow1 |> 
  fit(data = data_juice1)

saveRDS(fit_rf1, file = "data/report/fit_rf1.rds")
```

```{r}
# apply the model to test set
fit_rf1 <- readRDS("data/report/fit_rf1.rds")

va_ts_pred_rf <- bake(data_prep1, va_test) |>
  mutate(pviol = predict(fit_rf1,
                         bake(data_prep1, va_test))$.pred_class)
```

Variable importance of `at-the-spread` violations:

```{r}
# measure accuracy

cf_rf_ts1 <- va_ts_pred_rf |>
  count(Violation_1, pviol) |>
  group_by(Violation_1) |>
  mutate(Accuracy = n[Violation_1==pviol]/sum(n)) |>
  pivot_wider(names_from = "pviol", 
              values_from = n, values_fill = 0)


bal_accuracy_rf1 <- bal_accuracy(va_ts_pred_rf, Violation_1, pviol)$.estimate

rf_model1 <- pull_workflow_fit(fit_rf1)$fit
importance(rf_model1)
```

Confusion Matrix and Statistics of `at-the-spread` violations on the test dataset:

```{r}
confusionMatrix(va_ts_pred_rf$pviol, va_ts_pred_rf$Violation_1)
```

#### Outside the Spread

```{r}
data_rec2 <- recipe(Violation_2 ~ ., data = vo_train) |>
  step_downsample(Violation_2) |>
  step_rm(Instrument, Year, Month)
```

```{r}
# data_prep2 <- prep(data_rec2)
# saveRDS(data_prep2, file = "data/report/data_prep2.rds")

data_prep2 <- readRDS("data/report/data_prep2.rds")
data_juice2 <- juice(data_prep2)
```

```{r, eval = FALSE}
# tuning hyperparameters
rf_tune2 <- rand_forest(
  mtry = tune(),
  trees = 1000,
  min_n = tune()) |>
  set_mode("classification") |>
  set_engine("randomForest")

tune_wf2 <- workflow() |>
  add_recipe(data_rec2) |>
  add_model(rf_tune2)
  
doParallel::registerDoParallel() # for parallel processing

set.seed(3344)
w_folds2 <- vfold_cv(vo_train)

rf_grid2 <- grid_regular(mtry(range = c(2, 8)),
                        min_n(range = c(1, 4)),
                        levels = 3)

tune_res2 <- 
  tune_grid(tune_wf2,
            resamples = w_folds2,
            grid = rf_grid2
            )

# select best combination based on accuracy
best_hp2 <- tune_res2 |> select_best(metric = "accuracy")
```

```{r, eval = FALSE}
# fit the model
set.seed(3344)
rf2 <- rand_forest(mtry = 5, 
                     min_n = 4) |>
  set_mode("classification") |>
  set_engine("randomForest", importance = TRUE)

rf_workflow2 <- workflow() |>
  add_model(rf2) |>
  add_formula(Violation_2 ~ .)

fit_rf2 <- rf_workflow2 |> 
  fit(data = data_juice2)

saveRDS(fit_rf2, file = "data/report/fit_rf2.rds")
```

```{r}
# apply the model to test set
fit_rf2 <- readRDS("data/report/fit_rf2.rds")

vo_ts_pred_rf <- bake(data_prep2, vo_test) |>
  mutate(pviol = predict(fit_rf2,
                         bake(data_prep2, vo_test))$.pred_class)
```

Variable importance of `outside-the-spread` violation:

```{r}
# measure accuracy
cf_rf_ts2 <- vo_ts_pred_rf |>
  count(Violation_2, pviol) |>
  group_by(Violation_2) |>
  mutate(Accuracy = n[Violation_2==pviol]/sum(n)) |>
  pivot_wider(names_from = "pviol", 
              values_from = n, values_fill = 0)

bal_accuracy_rf2 <- bal_accuracy(vo_ts_pred_rf, Violation_2, pviol)$.estimate

rf_model2 <- pull_workflow_fit(fit_rf2)$fit
importance(rf_model2)
```

Confusion Matrix and Statistics of `outside-spread` violations on the test dataset:

```{r}
confusionMatrix(vo_ts_pred_rf$pviol, vo_ts_pred_rf$Violation_2)
```

#### Apply model to twpi data outside broker 210 and 361

```{r}
# apply at-spread violation model and measure accuracy 
data_2_va_pred_rf <- bake(data_prep1, data_2) |>
  mutate(pviol = predict(fit_rf1,
                         bake(data_prep1, data_2))$.pred_class)

cf_rf_data_2_1 <- data_2_va_pred_rf |>
  count(Violation_1, pviol) |>
  group_by(Violation_1) |>
  mutate(Accuracy = n[Violation_1==pviol]/sum(n)) |>
  pivot_wider(names_from = "pviol", 
              values_from = n, values_fill = 0)

bal_accuracy_rf_data2_1 <- bal_accuracy(data_2_va_pred_rf, Violation_1, pviol)$.estimate

```

Confusion Matrix and Statistics of `at-the-spread` violations on the new dataset:

```{r}
confusionMatrix(data_2_va_pred_rf$pviol, data_2_va_pred_rf$Violation_1)
```

```{r}
# apply outside violation model and measure accuracy 
data_2_vo_pred_rf <- bake(data_prep2, data_2) |>
  mutate(pviol = predict(fit_rf2,
                         bake(data_prep2, data_2))$.pred_class)

cf_rf_data_2_2 <- data_2_vo_pred_rf |>
  count(Violation_2, pviol) |>
  group_by(Violation_2) |>
  mutate(Accuracy = n[Violation_2==pviol]/sum(n)) |>
  pivot_wider(names_from = "pviol", 
              values_from = n, values_fill = 0)

bal_accuracy_rf_data2_2 <- bal_accuracy(data_2_vo_pred_rf, Violation_2, pviol)$.estimate

```

Confusion Matrix and Statistics of `outside-spread` violations on the new dataset:

```{r}
confusionMatrix(data_2_vo_pred_rf$pviol, data_2_vo_pred_rf$Violation_2)
```

### Summary

```{r}

summary <- data.frame(
  Violation = c("at-spread", "outside-spread"),
  GLM_top2 = c(bal_accuracy_logistic_va, bal_accuracy_logistic_vo),
  GLM_others = c(bal_accuracy_logistic_data2_1, bal_accuracy_logistic_data2_2),
  RF_top2 = c(bal_accuracy_rf1, bal_accuracy_rf2),
  RF_others = c(bal_accuracy_rf_data2_1, bal_accuracy_rf_data2_2)
)


summary |>
  kbl() |>
  kable_styling(full_width = FALSE, position = "left")
```

### Limitation

# Conclusion 500

## Potential Future Work

# References

ASIC Market Integrity Rules (Competition in Exchange Markets) 2011.

Australian Securities and Investments Commission. (20 May 2004). *Market demutualisation and privatisation: The Australian experience* [Speech]. Retrieved 24 October 2024, from https://asic.gov.au/about-asic/news-centre/speeches/market-demutualisation-and-privatisation-the-australian-experience/

Australian Securities and Investments Commission. (May 2014). *Report 394: Review of recent rule changes affecting dark liquidity* (PDF). Retrieved 24 October 2024, from https://download.asic.gov.au/media/1344596/rep394-published-19-May-2014.pdf

Australian Securities and Investments Commission. (May 2022). *Market integrity update: Issue 137*. Retrieved 24 October 2024, from https://asic.gov.au/about-asic/corporate-publications/newsletters/market-integrity-update/miu-issue-137-may-2022/

Australian Securities and Investments Commission. (3 February 2023). *Wilson's Advisory and Stockbroking Ltd ACN 010 529 665 pays $548,328 infringement notice* [Media release]. Retrieved 24 October 2024, from https://asic.gov.au/about-asic/news-centre/find-a-media-release/2023-releases/23-016mr-wilsons-advisory-and-stockbroking-ltd-acn-010-529-665-pays-548-328-infringement-notice/

Australian Securities Exchange. (n.d.). ASX regulatory framework. Retrieved 24 October 2024, from https://www.asx.com.au/about/regulation

Australian Securities Exchange. (n.d.). ASX story. Retrieved 24 October 2024, from https://www.asx.com.au/about/asx-story

Australian Securities Exchange. (n.d.). About Us. Retrieved 24 October 2024, from https://www.asx.com.au/about#:~:text=As%20an%20integrated%20exchange%2C%20ASX's,equities%20and%20fixed%20income%20markets.

Australian Securities Exchange. (19 February 2024). *ASX Trade: Markets Participant and Trading Schedule of Fees* (PDF). Retrieved 24 October 2024, from https://asxonline.com/content/dam/asxonline/public/documents/schedule-of-fees/asx-trade-markets-participant-and-trading-schedule-of-fees.pdf

Charles Lane Advisory Pty Ltd. (May 2014). *Report by Charles Lane Advisory: Review of recent rule changes affecting dark liquidity* (PDF). Attachment to ASIC Report 394: Review of recent rule changes affecting dark liquidity. Retrieved 24 October 2024, from https://download.asic.gov.au/media/1344590/rep394-attachment-published-19-May-2014.pdf

